```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "D:/202512-dev-seq-analyses/")

source("data_import.R")
source("PA_measurement.R")
source("TA_measurement.R")
source("OA_measurement.R")
```

```{r}
library("geomorph")
library("dplyr")
options(rgl.useNULL=TRUE)
library("rgl")
library("tidyverse")
library("gtools")
library(lubridate)
library(gridExtra)
library(factoextra)
library(ggplot2)
library(clustvarsel)
library(mclust)
library(ggpubr)
library(svglite)

#new package
# devtools::install_github("millacarmona/morphospace")
library(morphospace)
library(geomorph)
library(Morpho)
# devtools::install_github("MomX/Momocs")
library(Momocs)
library(magrittr)
library(rgl)
```


# Import data file


```{r}
# df <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds")
#df <- readRDS("D:/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds")

#files <- c("D:/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds", "D:/202603-dev-seq-analysis/data/202509-dev-seq-v6.rds")

files <- c("/Volumes/Yinghui/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds", "/Volumes/Yinghui/202603-dev-seq-analysis/data/202509-dev-seq-v6.rds")

df <- lapply(files, readRDS) %>% bind_rows()
```

Make sure the landmarks used is the same as the df_iso
```{r}
df_NoLA <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/df_NoLA.rds")
```

```{r}
df <- df %>%
  mutate(
    date_string = str_extract(basename(file), "^\\d{4}"),
    date_obj = ymd(paste0("2025", date_string))
  ) %>%
  select(-date_string)
```

```{r}
df_matched <- inner_join(
  df, 
  df_NoLA, 
  by = c("date_obj", "temp", "ind","biorep")
)
```

```{r}
# convert the lm to matrix for each file.
specs_list <- df_matched %>%
  group_by(temp, biorep, ind, plane, file) %>%
  group_map(~ {
    id <- .y$file
    # Create a sub-list for this specific identifier
    spec <- list(
      matrix = as.matrix(.x[, c("X", "Y")]),
      metadata = as.list(.y)
    )
    setNames(list(spec), id)
  }) %>%
  flatten()
```

```{r}
# add lm 13 in the TA plane
specs_list<-add_lm13(specs_list)
```


# GM

## OA
```{r}
oa_idx <- sapply(specs_list, function(x) {
  x$metadata$plane == "OA"
})

# Filter the list to keep only those specimens
oa_list <- specs_list[oa_idx]

# Extract just the 'matrix' part from every list element
# This creates a list of 2D matrices
oa_m <- lapply(oa_list, function(x) x$matrix)

# Combine them into a 3D Array
# Dimensions will be [Landmarks, Coordinates, Specimens]
# In your case: [8, 2, n]
oa_3d_array <- simplify2array(oa_m)

# (Optional) Assign specimen names to the 3rd dimension for tracking
# dimnames(oa_3d_array)[[3]] <- names(format_filtered)

# Check the final structure
dim(oa_3d_array)
```
### GPA
```{r}
gpaOA <- gpagen(oa_3d_array)
plotAllSpecimens(gpaOA$coords)
```

###CVA

```{r}
CVA_stage=CVA(gpaOA$coords, df_NoLA$stage, weighting = TRUE,
	plot = TRUE, rounds = 1000, cv = TRUE)
CVA_stage
stage <- as.factor(df_NoLA$stage)
plot(CVA_stage$CVscores[,1], CVA_stage$CVscores[,2], asp = 1,
     col = stage,
     xlab = "CV1", ylab = "CV2")
legend('topright', legend = levels(stage), col = 1:4, cex = 0.8, pch = 1)
```

### morphospace

```{r}
shapes <- gpaOA$coords

lks <- list(c(2,3), c(4,5), c(6,7))

# Generate morphospace
mspace(shapes, groups = stage, links = lks, cex.ldm = 5)
```

```{r}
data("tails")

shapes <- tails$shapes
spp <- tails$data$species
wf <- tails$links
phy <- tails$tree

# Generate morphospace
mspace(shapes, links = wf, cex.ldm = 5)

```
### between-group PCA
```{r}
bwgPCA_stage = groupPCA(gpaOA$coords, stage, rounds = 10000, cv = TRUE, weighting = FALSE)
plot(bwgPCA_stage$Scores[,1],bwgPCA_stage$Scores[,2], col=stage, asp=1, xlab="bwgPC 1",ylab="bwgPC 2")
legend('topright', legend = levels(stage), col = 1:4, cex = 0.8, pch = 1)
```

```{r}
bwgPCA_stage$groupdists
bwgPCA_stage$probs
```
Determine the landmarks that contributed the most through PC loading

for PC1:
```{r}
PCA_OA <- gm.prcomp(gpaOA$coords)
plot(PCA_OA,col = stage)
```

```{r}
oa_rotation <- PCA_OA$rotation
oa_sdev <- PCA_OA$sdev

# 2. Multiply each rotation column by its component's standard deviation
oa_loadings <- sweep(oa_rotation, 2, oa_sdev, FUN = "*")

# View the true loadings
head(oa_loadings)
```
```{r}
loadings_df <- as.data.frame(oa_loadings)

# Group X and Y components per landmark
landmark_loadings <- sapply(seq(1, nrow(oa_loadings), by = 2), function(i) {
  sqrt(oa_loadings[i, ]^2 + oa_loadings[i + 1, ]^2)
})

# Transpose so landmarks are rows and components are columns
landmark_loadings <- t(landmark_loadings)
rownames(landmark_loadings) <- paste0("Landmark_", 1:(nrow(oa_loadings)/2))

# Display overall magnitude loading per landmark
head(landmark_loadings)
```

scree plot
```{r}
# 1. Calculate individual and cumulative variance explained
if (!is.null(PCA_OA$d)) {
  var_explained <- (PCA_OA$d^2) / sum(PCA_OA$d^2)
} else {
  var_explained <- (PCA_OA$sdev^2) / sum(PCA_OA$sdev^2)
}

scree_df <- data.frame(
  PC = factor(paste0("PC", 1:length(var_explained)), levels = paste0("PC", 1:length(var_explained))),
  Variance = var_explained * 100,
  Cumulative = cumsum(var_explained * 100)
)

# 2. Plot with ggplot2
ggplot(scree_df, aes(x = PC)) +
  geom_bar(aes(y = Variance), stat = "identity", fill = "steelblue", alpha = 0.8) +
  geom_line(aes(y = Cumulative, group = 1), color = "firebrick", size = 1) +
  geom_point(aes(y = Cumulative), color = "firebrick", size = 2) +
  labs(
    title = "Scree Plot with Cumulative Variance",
    x = "Principal Component",
    y = "Variance Explained (%)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Allometric analysis
```{r}
ShapeANDsize <- cbind(gpaOA$coords, log(gpaOA$Csize))
PCASizeShape <- prcomp(ShapeANDsize)

plot(PCASizeShape$x[,1],PCASizeShape$x[,2],
     col= stage,
     xlab="RW in size-shape space 1",ylab="RW in size-shape space 2",asp = 1)
legend('topright', legend = levels(stage), col = 1:4, cex = 0.8, pch = 1)

```

Use a linear model to conduct MANOVA

```{r}
#oa_shape <- gpaOA$coords
#oa_size <- log(gpaOA$Csize)
# can refer back to lm part 3
#ShapeOnSize <- lm(oa_shape ~ oa_size)
#summary(manova(ShapeOnSize))
```


## TA
```{r}
ta_idx <- sapply(specs_list, function(x) {
  x$metadata$plane == "TA"
})

# Filter the list to keep only those specimens
ta_list <- specs_list[ta_idx]

# Extract just the 'matrix' part from every list element
# This creates a list of 2D matrices
ta_m <- lapply(ta_list, function(x) x$matrix)

# Combine them into a 3D Array
# Dimensions will be [Landmarks, Coordinates, Specimens]
# In your case: [8, 2, n]
ta_3d_array <- simplify2array(ta_m)

# (Optional) Assign specimen names to the 3rd dimension for tracking
# dimnames(oa_3d_array)[[3]] <- names(format_filtered)

# Check the final structure
dim(ta_3d_array)
```

### GPA
```{r}
gpaTA <- gpagen(ta_3d_array)
plotAllSpecimens(gpaTA$coords)
```

```{r}
CVA_stage=CVA(gpaTA$coords, df_NoLA$stage, weighting = TRUE,
	plot = TRUE, rounds = 1000, cv = TRUE)
CVA_stage
stage <- as.factor(df_NoLA$stage)
plot(CVA_stage$CVscores[,1], CVA_stage$CVscores[,2], asp = 1,
     col = stage,
     xlab = "CV1", ylab = "CV2")
legend('topright', legend = levels(stage), col = 1:4, cex = 0.8, pch = 1)
```



```{r}
PCA_TA <- gm.prcomp(gpaTA$coords)
plot(PCA_TA,col = stage)
```

Determine the lanmark that contributed te most through PC loadings

for PC1:
```{r}

```

### between-group PCA
```{r}
bwgPCA_stage=groupPCA(gpaTA$coords, stage, rounds = 10000, cv = TRUE, weighting = FALSE)
plot(bwgPCA_stage$Scores[,1],bwgPCA_stage$Scores[,2], col=stage, asp=1, xlab="bwgPC 1",ylab="bwgPC 2")
legend('topright', legend = levels(stage), col = 1:4, cex = 0.8, pch = 1)
```

Allometric analysis
```{r}
ShapeANDsize <- cbind(gpaTA$coords, log(gpaTA$Csize))
PCASizeShape <- prcomp(ShapeANDsize)

plot(PCASizeShape$x[,1],PCASizeShape$x[,2],
     col= stage,
     xlab="RW in size-shape space 1",ylab="RW in size-shape space 2",asp = 1)
legend('topright', legend = levels(stage), col = 1:4, cex = 0.8, pch = 1)

```
### TA without lobe landmarks
```{r}
ta_final_points <- ta_3d_array[c(13,8,9,3,4,5,10,11,12,1,2) , ,]
gpaTA_2 <- gpagen(ta_final_points)

```
### morphospace

```{r}
shapes <- gpaTA_2$coords
#stage <- df_NoLA$stage
#lks <- list(c(13,3,4), c(10,5), c(8,9), c(5,3))
lks <- list(c(1,4,5), c(6,7), c(2,3), c(4,6))

# Generate morphospace
mspace(shapes, groups = stage, links = lks, cex.ldm = 5)
```

```{r}
# Get mean shapes of each stage
stg_shapes <- expected_shapes(shapes = shapes, x = stage)

# Generate morphospace and project:
msp <- mspace(shapes = shapes, links = lks, cex.ldm = 5) %>% 
  # scatter points
  proj_shapes(shapes = shapes, col = stage) %>% 
  # convex hulls enclosing groups
  proj_groups(shapes = shapes, groups = stage, alpha = 0.5) 
```
Allometric analysis
```{r}
# remove allometric variation for each species
d_out <- out * 0
for(i in 1:nlevels(sp)) {
  index <- which(sp == levels(sp)[i])
  mod <- lm(out[index,] ~ size[index])
  d_out[index,] <- detrend_shapes(mod,
                                  xvalue = max(size[index]),
                                  method = "residuals")
}

# get bgPCA
bgpca <- bg_prcomp(x = d_out, groups = sp)

# generate bgPCA morphospace with detrended shapes; project shapes and ellipses
mspace(ord = bgpca, datype = "fcoef", nh = 8, nv = 6, size.models = 1.5,
       bg.model = "light gray", invax = c(1,2)) %>%
  proj_shapes(shapes = d_out, col = sp) %>%
  proj_groups(shapes = d_out, groups = sp, alpha = 0.4, ellipse = TRUE)

```

## PA
```{r}
pa_idx <- sapply(specs_list, function(x) {
  x$metadata$plane == "PA"
})

# Filter the list to keep only those specimens
pa_list <- specs_list[pa_idx]

# Extract just the 'matrix' part from every list element
# This creates a list of 2D matrices
pa_m <- lapply(pa_list, function(x) x$matrix)

# Combine them into a 3D Array
# Dimensions will be [Landmarks, Coordinates, Specimens]
# In your case: [8, 2, n]
pa_3d_array <- simplify2array(pa_m)

# (Optional) Assign specimen names to the 3rd dimension for tracking
# dimnames(oa_3d_array)[[3]] <- names(format_filtered)

# Check the final structure
dim(pa_3d_array)
```

### GPA
```{r}
gpaPA <- gpagen(pa_3d_array)
plotAllSpecimens(gpaPA$coords)
```
### LDA

```{r}
CVA_stage=CVA(gpaPA$coords, df_NoLA$stage, weighting = TRUE,
	plot = TRUE, rounds = 1000, cv = TRUE)
CVA_stage
stage <- as.factor(df_NoLA$stage)
plot(CVA_stage$CVscores[,1], CVA_stage$CVscores[,2], asp = 1,
     col = stage,
     xlab = "CV1", ylab = "CV2")
legend('topright', legend = levels(stage), col = 1:4, cex = 0.8, pch = 1)
```
### between-group PCA
```{r}
bwgPCA_stage=groupPCA(gpaPA$coords, stage, rounds = 10000, cv = TRUE, weighting = FALSE)
plot(bwgPCA_stage$Scores[,1],bwgPCA_stage$Scores[,2], col=stage, asp=1, xlab="bwgPC 1",ylab="bwgPC 2")
legend('topright', legend = levels(stage), col = 1:4, cex = 0.8, pch = 1)
```
