---
title: "202512_morph_lm"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "D:/202512-dev-seq-analyses/")

source("data_import.R")
source("PA_measurement.R")
source("TA_measurement.R")
source("OA_measurement.R")
```

```{r}
library("geomorph")
library("dplyr")
options(rgl.useNULL=TRUE)
library("rgl")
library("tidyverse")
library("gtools")
library(lubridate)
library(gridExtra)
library(factoextra)
library(ggplot2)
library(clustvarsel)
library(mclust)
library(ggpubr)
library(svglite)
```


# Import data file

Dir:"/Volumes/Yinghui/202512-dev-seq-analyses"

```{r}
# df <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds")
#df <- readRDS("D:/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds")

#files <- c("D:/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds", "D:/202603-dev-seq-analysis/data/202509-dev-seq-v6.rds")

files <- c("/Volumes/Yinghui/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds", "/Volumes/Yinghui/202603-dev-seq-analysis/data/202509-dev-seq-v6.rds")

df <- lapply(files, readRDS) %>% bind_rows()
```



```{r}
# convert the lm to matrix for each file.

specs_list <- df %>%
  group_by(temp, biorep, ind, plane, file) %>%
  group_map(~ {
    id <- .y$file
    # Create a sub-list for this specific identifier
    spec <- list(
      matrix = as.matrix(.x[, c("X", "Y")]),
      metadata = as.list(.y)
    )
    
    setNames(list(spec), id)
  }) %>%
  flatten()

```

```{r}
# add lm 13 in the TA plane
specs_list<-add_lm13(specs_list)
```


# Plot setting
```{r}
palette1 <- c("#343434","#A3CEF1","#3880ba","#224f73","#000000","#CCCAC2","#837D72","#ede7da")
palette2 = c("#f1eef6","#d0d1e6","#a6bddb","#74a9cf","#3690c0",
"#0570b0","#034e7b")
palette3 = c("#E8F0F8FF", "#D8D8E8FF", "#C0C0D0FF", "#70B0E8FF", "#6088C0FF", "#406898FF", "#283858FF", "#000000FF", "#A8B0C0FF", "#808090FF", "#A8D0F8FF", "#802008FF")
```

#linear measurements
```{r}
measurement_accumulator <- list()

for (id in names(specs_list)) {
  
  # 1. Extract components correctly
  spec   <- specs_list[[id]]
  mat    <- spec$matrix
  plane  <- spec$metadata$plane
  ind    <- spec$metadata$ind
  temp <- spec$metadata$temp
  biorep <- spec$metadata$biorep
  file <- spec$metadata$file
  measurer <- spec$metadata$measurer
  
 # print(file)
  # 2. Initialize a placeholder for this specific iteration's results
  res <- tibble()
  
 # print(file)
  # 3. Apply custom "Design" logic
  if (plane == "PA") {
    res <- tibble(
      measure_name = c("ALO", "AW", "WAE", "ATCL", "PABW"),
      value = c(ALO_fun(mat), AW_fun(mat), WAE_fun(mat), ATCL_fun(mat), PABW_fun(mat))
    )
    
  } else if (plane == "TA") {
    res <- tibble(
      measure_name = c("PL", "BL", "TCL", "LTM", "LL", "TABW", "AECL","TATB"),
      value = c(PL_fun(mat), BL_fun(mat), TCL_fun(mat), LTM_fun(mat), 
                LL_fun(mat), TABW_fun(mat), AECL_fun(mat), TATB_fun(mat))
    )
    
  } else if (plane == "OA") {
    res <- tibble(
      measure_name = c("TBL", "PW", "ATAC", "AEAC"),
      value = c(TBL_fun(mat), PW_fun(mat), ATAC_fun(mat), AEAC_fun(mat))
    )
  }
  
  # print(file)
  # print(row(mat))
  # print(file)
  # 4. Add metadata and store in accumulator
  # Only add if we actually generated results
  if (nrow(res) > 0) {
    measurement_accumulator[[id]] <- res %>%
      mutate(
        measurer = measurer,
        sample_id = str_extract(basename(file), "^\\d+"),
        ind = ind,
        plane = plane,
        temp = temp,
        biorep = biorep
      )
  }
}
# 5. Bind into one large dataframe
df_lm <- bind_rows(measurement_accumulator)

df_lm <- df_lm %>% mutate(
    # Convert from microns (or current unit) to mm
    value = value / 1000
  ) 
```

Removing the value of lobate and auricle when those structures are not present
standard:
ALO 0.07mm
AW 0.07mm
LTM 0.07mm
LL 0.07mm
```{r}
threshold <- 0.07

df_lm <- df_lm %>% mutate(across(
    c(value), 
    ~ if_else(measure_name %in% c("ALO", "AW", "LTM", "LL") & .x < threshold, NA_real_, .x)
  ))
  
```

convert the sample_id to date
```{r}
df_lm <- df_lm %>%
  mutate(date_obj = parse_date_time(paste0(sample_id, "2026"), orders = "mdy")) %>%
  group_by(biorep) %>%
  mutate(exp_day = as.numeric(difftime(date_obj, min(date_obj), units = "days"))) %>%
  ungroup() #%>%
 # dplyr::select(-c(date_obj))
```
change the sample id to individual
```{r}
df_lm <- df_lm  %>%
  mutate(sample_id = paste(ind, temp, biorep, sep = "-"))
```

## allometric dev visualisation

### obtain the natural log of the value
Common practice when studying allometric development because it is easier to deal with proportional variation.

```{r}
#create a new column with log value
df_lm <- df_lm %>% mutate(log_value = log(value))
```

###Plotting for the shape versus size LMM study

centroid size is recalculated as a Solid semi-ellipsoid around z-axis (without lobe)

volume: (2*pi*a*b*c)/3

a, b = TABW/2, PABW/2, the principal semi-axes of the base ellipse
c = BL,the principal z-semi-axe from the center of base ellipse

```{r}
df_lmw <- df_lm %>%
  dplyr::select(-c(plane, log_value)) %>%
  group_by(sample_id, ind) %>%
  pivot_wider(
    names_from = measure_name,
    values_from = value
  ) 

df_lmw <- df_lmw %>% 
  mutate(centroid = (2 *(TABW/2) * (PABW/2)*BL )/3)


measure_cols <- colnames(df_lmw[,6:22])
df_iso <-  df_lmw  %>%  
  mutate(across(all_of(measure_cols), 
                ~ log(.x/centroid), 
                .names = "shape_{.col}"))

```


```{r}
p1 <-ggplot(df_lmw, aes(x = log(centroid), y = log(TBL), colour = temp)) +
    geom_point(size = 2, alpha = 0.8) +
    # Add the dashed regression line
    geom_smooth(method = "lm", formula = y ~ x, color = "black") +
    theme_classic() +
    labs(x = "Log Centroid size", y = "Log of the tentacle bulb length (mm)") +
    scale_color_manual(values = c( "#00B2CA", "#F79256")) +
    theme(legend.position = "right")
#+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))
p4 <-ggplot(df_lmw, aes(x = log(centroid), y = log(TABW), colour = temp)) +
    geom_point(size = 2, alpha = 0.8) +
    # Add the dashed regression line
    geom_smooth(method = "lm", formula = y ~ x, color = "black") +
    theme_classic() +
    labs(x = "Log Centroid size", y = "Log of the body width (mm)") +
    scale_color_manual(values = c( "#00B2CA", "#F79256")) +
    theme(legend.position = "right")
#+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))
p2 <- ggplot(df_lmw, aes(x = log(centroid), y = log(LL), colour = temp)) +
    geom_point(size = 2, alpha = 0.8) +
    # Add the dashed regression line
    geom_smooth(method = "lm", formula = y ~ x, color = "black") +
    theme_classic() +
    labs(x = "Log Centroid size", y = "Log of the lobe side length (mm)") +
    scale_color_manual(values = c("#00B2CA", "#F79256")) +
    theme(legend.position = "right")
#+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))
p3 <- ggplot(df_lmw, aes(x = log(centroid), y = log(ALO), colour = temp)) +
    geom_point(size = 2, alpha = 0.8) +
    # Add the dashed regression line
    geom_smooth(method = "lm", formula = y ~ x, color = "black") +
    theme_classic() +
    labs(x = "Log Centroid size", y = "Log of the auricle length (mm)") +
    scale_color_manual(values = c("#00B2CA", "#F79256")) +
    theme(legend.position = "right")
#+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))

p <- ggarrange(p1,p4, p2,p3, nrow = 2, ncol = 2,common.legend = TRUE)
#ggsave(file="/Volumes/Yinghui/202512-dev-seq-analyses/fig/p.svg", plot=p)
#ggsave(file="D:/202603-dev-seq-analysis/fig/p1.svg", plot=p)

```

#### With all variables:

```{r}
measure_cols <- colnames(df_lmw[,7:23])
df_iso <-  df_lmw  %>%  
  mutate(across(all_of(measure_cols), 
                ~ log(.x/centroid), 
                .names = "shape_{.col}"))
```

```{r}
library("RRPP")
# df_clean <- df_iso %>% filter(across(where(is.numeric), is.finite)) %>% ungroup()

# Create a matrix of your shape variables
shape_matrix <- df_iso %>% ungroup()%>%
  dplyr::select(-c("LL","LTM","ALO","AW"))%>%
  dplyr::select(starts_with("shape_")) %>%
  drop_na() %>%
   filter(across(where(is.numeric), is.finite)) %>% 
  as.matrix()


df_temp <- df_iso %>% ungroup()%>%
  dplyr::select(-c("LL","LTM","ALO","AW"))%>%
  drop_na()%>%
   filter(across(where(is.numeric), is.finite))

#storage.mode(shape_matrix) <- "double"

nrow(shape_matrix)
nrow(df_temp)

# Add the centroid back to a clean modeling data frame
model_data <- rrpp.data.frame(
  centroid = df_temp$centroid,
  coords = shape_matrix,
  temp = as.factor(df_temp$temp)
)
#any(!is.finite(shape_matrix)) # Should be FALSE
#table(model_data$temp)
fit <- lm.rrpp(coords ~ log(centroid)*temp, SS.type = "I", 
data = model_data, print.progress = FALSE, iter = 999)
summary(fit)
anova(fit)
```
## GMM

if only considering shape (because there are temperature as factor, so measurements not normalized by size / body length is not appropriate)
```{r}
df_NoLA <- df_iso %>% 
  dplyr::select(-c("ALO", "AW", "LTM", "LL","shape_LL","shape_LTM","shape_ALO","shape_AW"))%>%
  drop_na()%>%
  filter(across(where(is.numeric), is.finite))

df_ori <- df_NoLA[,7:19]
df_log <- log(df_NoLA[,7:19])
df_log_shape_ratio <- df_NoLA[,21:33]
mod1 <- Mclust(df_ori,G=1:12)
summary(mod1)
mod2 <- Mclust(df_log,G=1:12)
summary(mod2)
mod3 <- Mclust(df_log_shape_ratio,G=1:12)
summary(mod3)
```
```{r}
p <- fviz_mclust(mod1,"classification",  geom = "point" ) +
#scale_shape_manual(values = c(16,0,17,8))
scale_color_manual(values = palette3)+
scale_fill_manual( values = palette3)
p2 <- fviz_mclust(mod2,"classification",  geom = "point" ) +
#scale_shape_manual(values = c(16,0,17,8))
scale_color_manual(values = palette3)+
scale_fill_manual( values = palette3)
p3 <- fviz_mclust(mod3,"classification",  geom = "point" ) +
#scale_shape_manual(values = c(16,0,17,8))
scale_color_manual(values = palette3)+
scale_fill_manual( values = palette3)
pGMM <- ggarrange(p, p2, p3, nrow = 2, ncol = 2,common.legend = TRUE)
#ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/tatb_vs_centroid.svg", plot=pp1)
#ggsave(file="D:/202603-dev-seq-analysis/fig/GMM_PCA.svg", plot=p)
```

```{r}
#ggsave(file="D:/202603-dev-seq-analysis/fig/GMM_PCA.svg", plot=p)

pca <- prcomp(scale(mod1$data))
fviz_contrib(pca, choice = "var", axes = 1, top = 10)
fviz_contrib(pca, choice = "var", axes = 2, top = 10)
fviz_contrib(pca, choice = "var", axes = 3, top = 10)

```

construct the plot after selecting the variables
```{r}
#out1 <- clustvarsel(df_ori)
#mod1_1 <-Mclust(df_ori[, out1$subset])
#summary(mod1_1)
#fviz_mclust(mod1_1,"classification",  geom = "point" ) 

```

```{r}
library(ggalluvial)
df_iso <- df_iso %>%
mutate(stage = case_when(
is.na(ALO) & is.na(LL)  ~ "cydippid",
is.na(ALO) & !is.na(LL) ~ "early_transition",
!is.na(ALO) & !is.na(LL) ~ "late_transition",
TRUE                     ~ "unclassified"
))

df_NoLA <- df_iso %>% ungroup()%>%
dplyr::select(-c("ALO", "AW", "LTM", "LL","shape_LL","shape_LTM","shape_ALO","shape_AW"))%>%
drop_na()%>%
filter(across(where(is.numeric), is.finite))

df_NoLA$cluster <- as.factor(mod3$classification)
df_NoLA$cluster_logsize <- as.factor(mod2$classification)
df_NoLA$cluster_size <- as.factor(mod1$classification)

ggplot(df_NoLA, aes(x = log(centroid), y = log(TABW), color = cluster)) +
geom_point(size = 1) +
theme_classic()
#scale_color_manual(values = c("#343434","#ccc9c1","#a4cdee","#3880ba"))
```

```{r}
g_plot <- ggplot(df_NoLA,
aes(axis1 = cluster, axis2 = stage, y = ..count..)) +
geom_alluvium(aes(fill = cluster)) +
geom_stratum() +
theme_classic() +
labs(x = "Classification Method",
y = "Number of images")
#scale_fill_manual(values = c("#343434","#3880ba","#a4cdee"))
g_plot
#ggsave(file="D:/202603-dev-seq-analysis/fig/ggalluvial.svg", plot=g_plot)
```

```{r}
ggplot(df_NoLA, aes(x = log(centroid), y = log(TABW), color = cluster)) +
labs(x = "Log of body length (mm)",y = "log of body width (mm)")+
geom_point(size = 1) +
theme_classic()
#scale_color_manual(values = c("#343434","#ccc9c1","#a4cdee","#3880ba"))
```
## segmented
```{r}
library(segmented)
measure_cols <- colnames(df_NoLA[,7:20])
df_NoLA <- df_NoLA %>%
  mutate(
    across(all_of(measure_cols), 
                ~ log(.x), 
                .names = "log_{.col}")
  )%>%
  mutate(log_centroid = log(centroid))

model1 <- lm(log_TBL ~ log_centroid, data = df_NoLA)

segmented_model1 <- segmented(model1, seg.Z = ~log_centroid)

df_NoLA$seg_preds1 <- predict(segmented_model1)

pp1 <- ggplot(df_NoLA, aes(x = log_centroid, y = log_TBL, color = cluster)) +
  #geom_point(color = "#3880ba", alpha = 0.8) +                # Scatter plot of raw data
  geom_line(aes(y = seg_preds1), color = "#343434", size = 1) + # Piecewise regression line
  theme_classic() +
  geom_point()+
    labs(y = "Log TBL")
```

```{r}
model1 <- lm(log_TATB ~ log_centroid, data = df_NoLA)

segmented_model1 <- segmented(model1, seg.Z = ~log_centroid)

df_NoLA$seg_preds1 <- predict(segmented_model1)

ppTATB <- ggplot(df_NoLA, aes(x = log_centroid, y = log_TATB, color = stage)) +
  #geom_point(color = "#3880ba", alpha = 0.8) +                # Scatter plot of raw data
  geom_line(aes(y = seg_preds1), color = "#343434", size = 1) + # Piecewise regression line
  theme_classic() +
  geom_point()+
    labs(y = "Log TATB")
#ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/tatb_vs_centroid.svg", plot=pp1)
```


```{r}
ggplot(df_NoLA, aes(x = TATB , y = TBL, color = stage)) +
  #geom_point(color = "#3880ba", alpha = 0.8) +                # Scatter plot of raw data
  #geom_line(aes(y = seg_preds1), color = "#343434", size = 1) + # Piecewise regression line
  theme_classic() +
  geom_point()
```

```{r}
df_iso<-  df_iso %>%
mutate(
log_AW = log(AW),
log_centroid = log(centroid)
)
df_temp <- df_iso %>% drop_na()
model2 <- lm(log_AW ~ log_centroid, data =  df_temp)
segmented_model2 <- segmented(model2, seg.Z = ~log_centroid)
df_temp$seg_preds2 <- predict(segmented_model2)
pp2 <- ggplot( df_temp, aes(x = log_centroid, y = log_AW, color = exp_day)) +
#geom_point(color = "#3880ba", alpha = 0.8) +                # Scatter plot of raw data
geom_line(aes(y = seg_preds2), color = "#343434", size = 1) + # Piecewise regression line
theme_classic() +
geom_point() +
scale_colour_gradient(high = "#132B43", low = "#56B1F7")+
    labs(y = "Log AW")
#ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/AW_vs_centroid.svg", plot=pp2)
#ggsave(file="D:/202603-dev-seq-analysis/fig/AW_vs_centroid.svg", plot=pp2)
```

```{r}
df_iso<-  df_iso %>%
mutate(
log_LL = log(LL)
)
df_temp <- df_iso %>% drop_na()
model2 <- lm(log_LL ~ log_centroid, data =  df_temp)
segmented_model2 <- segmented(model2, seg.Z = ~log_centroid)
df_temp$seg_preds2 <- predict(segmented_model2)
ppLL <- ggplot( df_temp, aes(x = log_centroid, y = log_LL, color = exp_day)) +
#geom_point(color = "#3880ba", alpha = 0.8) +                # Scatter plot of raw data
geom_line(aes(y = seg_preds2), color = "#343434", size = 1) + # Piecewise regression line
theme_classic() +
geom_point() +
scale_colour_gradient(high = "#132B43", low = "#56B1F7")+
    labs(y = "Log LL")
#ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/AW_vs_centroid.svg", plot=pp2)
#ggsave(file="D:/202603-dev-seq-analysis/fig/AW_vs_centroid.svg", plot=pp2)
```

```{r}
df_NoLA <- df_NoLA %>%
mutate(
log_TCL = log(TCL),
)
model1 <- lm(log_TCL ~ log_centroid, data = df_NoLA)
segmented_model1 <- segmented(model1, seg.Z = ~log_centroid)
# getting slope and breakpoint
segmented_model1$psi
slope(segmented_model1)
df_NoLA$seg_preds1 <- predict(segmented_model1)
pp3 <- ggplot(df_NoLA, aes(x = log_centroid, y = log_TCL, color = stage)) +
#geom_point(color = c("#A3CEF1","#3880ba","#343434"), alpha = 0.8) +                # Scatter plot of raw data
geom_point()+
  geom_line(aes(y = seg_preds1), color = "#343434", size = 1) + # Piecewise regression line
theme_classic() +
scale_color_manual(values = c("cydippid" = "#a4cdee",
"early_transition" = "#3880ba",
"late_transition" = "#343434"))+
    labs(y = "Log TCL")
#ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/tatb_vs_centroid.svg", plot=pp1)
#ggsave(file="D:/202603-dev-seq-analysis/fig/tatb_vs_centroid.svg", plot=pp1)
```

```{r}
df_NoLA <- df_NoLA %>%
mutate(
log_AECL = log(AECL),
log_ATCL = log(ATCL)
)
model4 <- lm(log_AECL ~ log_ATCL, data = df_NoLA)
segmented_model4 <- segmented(model4, seg.Z = ~log_ATCL)
# getting slope and breakpoint
segmented_model4$psi
slope(segmented_model4)
df_NoLA$seg_preds4 <- predict(segmented_model4)
pp4 <- ggplot(df_NoLA, aes(x = log_ATCL, y = log_AECL, color = stage)) +
#geom_point(color = c("#A3CEF1","#3880ba","#343434"), alpha = 0.8) +                # Scatter plot of raw data
geom_line(aes(y = seg_preds4), color = "#343434", size = 1) + # Piecewise regression line
theme_classic() +
geom_point()+
scale_color_manual(values = c("cydippid" = "#a4cdee",
"early_transition" = "#3880ba",
"late_transition" = "#343434"))+
    labs(y = "Log AECL")
#ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/tatb_vs_centroid.svg", plot=pp1)
#ggsave(file="D:/202603-dev-seq-analysis/fig/tatb_vs_centroid.svg", plot=pp1)
```

```{r}
df_NoLA <- df_NoLA %>%
mutate(
log_TABW = log(TABW),
log_BL = log(BL)
)
model4 <- lm(log_TABW ~ log_BL, data = df_NoLA)
segmented_model4 <- segmented(model4, seg.Z = ~log_BL)
# getting slope and breakpoint
segmented_model4$psi
slope(segmented_model4)
df_NoLA$seg_preds4 <- predict(segmented_model4)
pp5 <- ggplot(df_NoLA, aes(x = log_BL, y = log_TABW, color = stage)) +
#geom_point(color = c("#A3CEF1","#3880ba","#343434"), alpha = 0.8) +                # Scatter plot of raw data
geom_point(aes(alpha = 0.6))+
  geom_line(aes(y = seg_preds4), color = "#343434", size = 1) + # Piecewise regression line
theme_classic() +
scale_color_manual(values = c("cydippid" = "#a4cdee",
"early_transition" = "#3880ba",
"late_transition" = "#343434"))+
    labs(y = "Log TABW")
#ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/tatb_vs_centroid.svg", plot=pp1)
#ggsave(file="D:/202603-dev-seq-analysis/fig/tatb_vs_centroid.svg", plot=pp1)
```

```{r}
model4 <- lm(log_ATAC ~ log_centroid, data = df_NoLA)
segmented_model4 <- segmented(model4, seg.Z = ~log_centroid)
# getting slope and breakpoint
segmented_model4$psi
slope(segmented_model4)
df_NoLA$seg_preds4 <- predict(segmented_model4)
pp6 <- ggplot(df_NoLA, aes(x = log_centroid, y = log_ATAC, color = stage)) +
#geom_point(color = c("#A3CEF1","#3880ba","#343434"), alpha = 0.8) +                # Scatter plot of raw data
  geom_point(aes(alpha = 0.6))+
geom_line(aes(y = seg_preds4), color = "#343434", size = 1) + # Piecewise regression line
theme_classic() +
scale_color_manual(values = c("cydippid" = "#a4cdee",
"early_transition" = "#3880ba",
"late_transition" = "#343434"))+
    labs(y = "Log ATAC")
#ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/tatb_vs_centroid.svg", plot=pp1)
#ggsave(file="D:/202603-dev-seq-analysis/fig/tatb_vs_centroid.svg", plot=pp1)
```

Put together a figure
```{r}
#ggsave(
#   file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/stage_strcture_1.pdf", 
   plot = ggarrange(pp1,ppTATB, pp2, ppLL, pp3,pp4, nrow = 3, ncol=2)
plot
#)
```

```{r}
#ggsave(
#   file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/stage_strcture_2.pdf", 
   plot = ggarrange(pp5, pp6, nrow = 1, ncol=2)
#)
plot
```

```{r}
#ggsave(
#   file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/cluster_2.pdf", 
   plot = ggarrange(p, p2, p3, nrow = 2, ncol=2)
#)
plot
```

# 0417 new added
```{r}
model4 <- lm(log_ATAC ~ log_centroid, data = df_NoLA)
segmented_model4 <- segmented(model4, seg.Z = ~log_centroid)
# getting slope and breakpoint
segmented_model4$psi
slope(segmented_model4)
df_NoLA$seg_preds4 <- predict(segmented_model4)
pp7 <- ggplot(df_NoLA, aes(x = log_centroid, y = log_ATAC, color = cluster)) +
#geom_point(color = c("#A3CEF1","#3880ba","#343434"), alpha = 0.8) +                # Scatter plot of raw data
  geom_point()+
geom_line(aes(y = seg_preds4), color = "#343434", size = 1) + # Piecewise regression line
theme_classic() +
    labs(y = "Log ATAC")
#ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/tatb_vs_centroid.svg", plot=pp1)
#ggsave(file="D:/202603-dev-seq-analysis/fig/tatb_vs_centroid.svg", plot=pp1)
```

```{r}
model1 <- lm(log_TCL ~ log_centroid, data = df_NoLA)
segmented_model1 <- segmented(model1, seg.Z = ~log_centroid)
# getting slope and breakpoint
segmented_model1$psi
slope(segmented_model1)
df_NoLA$seg_preds1 <- predict(segmented_model1)
pp8 <- ggplot(df_NoLA, aes(x = log_centroid, y = log_TCL, color = cluster)) +
#geom_point(color = c("#A3CEF1","#3880ba","#343434"), alpha = 0.8) +                # Scatter 
  geom_point()+
geom_line(aes(y = seg_preds1), color = "#343434", size = 1) + # Piecewise regression line
theme_classic() +
    labs(y = "Log TCL")
```

```{r}
#ggsave(
#   file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/stage_cluster_1.pdf", 
   plot = ggarrange(pp6, pp7, pp3, pp8, nrow = 2, ncol=2)
#)
plot
```