---
title: "202512_morph_lm"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "D:/202512-dev-seq-analyses/")

source("data_import.R")
source("PA_measurement.R")
source("TA_measurement.R")
source("OA_measurement.R")
```

```{r}
library("geomorph")
library("dplyr")
options(rgl.useNULL=TRUE)
library("rgl")
library("tidyverse")
library("gtools")
library(lubridate)
library(gridExtra)
library(factoextra)
library(ggplot2)
library(clustvarsel)
library(mclust)
library(ggpubr)
library(svglite)
```


# Import data file


```{r}
# df <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds")
#df <- readRDS("D:/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds")

#files <- c("D:/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds", "D:/202603-dev-seq-analysis/data/202509-dev-seq-v6.rds")

files <- c("/Volumes/Yinghui/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds", "/Volumes/Yinghui/202603-dev-seq-analysis/data/202509-dev-seq-v6.rds")

df <- lapply(files, readRDS) %>% bind_rows()
```



```{r}
# convert the lm to matrix for each file.

specs_list <- df %>%
  group_by(temp, biorep, ind, plane, file) %>%
  group_map(~ {
    id <- .y$file
    # Create a sub-list for this specific identifier
    spec <- list(
      matrix = as.matrix(.x[, c("X", "Y")]),
      metadata = as.list(.y)
    )
    
    setNames(list(spec), id)
  }) %>%
  flatten()

```

```{r}
# add lm 13 in the TA plane
specs_list<-add_lm13(specs_list)
```


# Plot setting
```{r}
palette1 <- c("#343434","#A3CEF1","#3880ba","#224f73","#000000","#CCCAC2","#837D72","#ede7da")
palette2 = c("#f1eef6","#d0d1e6","#a6bddb","#74a9cf","#3690c0",
"#0570b0","#034e7b")
palette3 = c("#E8F0F8FF", "#D8D8E8FF", "#C0C0D0FF", "#70B0E8FF", "#6088C0FF", "#406898FF", "#283858FF", "#000000FF", "#A8B0C0FF", "#808090FF", "#A8D0F8FF", "#802008FF")
```

#linear measurements
```{r}
measurement_accumulator <- list()

for (id in names(specs_list)) {
  
  # 1. Extract components correctly
  spec   <- specs_list[[id]]
  mat    <- spec$matrix
  plane  <- spec$metadata$plane
  ind    <- spec$metadata$ind
  temp <- spec$metadata$temp
  biorep <- spec$metadata$biorep
  file <- spec$metadata$file
  measurer <- spec$metadata$measurer
  
 # print(file)
  # 2. Initialize a placeholder for this specific iteration's results
  res <- tibble()
  
 # print(file)
  # 3. Apply custom "Design" logic
  if (plane == "PA") {
    res <- tibble(
      measure_name = c("ALO", "AW", "WAE", "ATCL", "PABW"),
      value = c(ALO_fun(mat), AW_fun(mat), WAE_fun(mat), ATCL_fun(mat), PABW_fun(mat))
    )
    
  } else if (plane == "TA") {
    res <- tibble(
      measure_name = c("PL", "BL", "TCL", "LTM", "LL", "TABW", "AECL","TATB"),
      value = c(PL_fun(mat), BL_fun(mat), TCL_fun(mat), LTM_fun(mat), 
                LL_fun(mat), TABW_fun(mat), AECL_fun(mat), TATB_fun(mat))
    )
    
  } else if (plane == "OA") {
    res <- tibble(
      measure_name = c("TBL", "PW", "ATAC", "AEAC"),
      value = c(TBL_fun(mat), PW_fun(mat), ATAC_fun(mat), AEAC_fun(mat))
    )
  }
  
  # print(file)
  # print(row(mat))
  # print(file)
  # 4. Add metadata and store in accumulator
  # Only add if we actually generated results
  if (nrow(res) > 0) {
    measurement_accumulator[[id]] <- res %>%
      mutate(
        measurer = measurer,
        sample_id = str_extract(basename(file), "^\\d+"),
        ind = ind,
        plane = plane,
        temp = temp,
        biorep = biorep
      )
  }
}
# 5. Bind into one large dataframe
df_lm <- bind_rows(measurement_accumulator)

df_lm <- df_lm %>% mutate(
    # Convert from microns (or current unit) to mm
    value = value / 1000
  ) 
```

Removing the value of lobate and auricle when those structures are not present
standard:
ALO 0.07mm
AW 0.07mm
LTM 0.05mm
LL 0.07mm
```{r}
df_lm <- df_lm %>% 
  mutate(value = case_when(
    # Rule for LTM (Threshold 0.05)
    measure_name == "LTM"  & value < 0.05 ~ NA_real_,
    
    # Rule for ALO, AW, LL (Threshold 0.07)
    measure_name %in% c("ALO", "AW", "LL") & value < 0.07 ~ NA_real_,
    
    # Keep everything else as is
    TRUE ~ value
  ))
```

convert the sample_id to date
```{r}
df_lm <- df_lm %>%
  mutate(date_obj = parse_date_time(paste0(sample_id, "2025"), orders = "mdy")) %>%
  group_by(biorep) %>%
  mutate(exp_day = as.numeric(difftime(date_obj, min(date_obj), units = "days")+1)) %>%
  ungroup() 
 # dplyr::select(-c(date_obj))
```

change the sample id to individual
```{r}
df_lm <- df_lm  %>%
  mutate(sample_id = paste(ind, temp, biorep, sep = "-"))
```

## allometric dev visualisation

### obtain the natural log of the value
Common practice when studying allometric development because it is easier to deal with proportional variation.

```{r}
#create a new column with log value
df_lm <- df_lm %>% mutate(log_value = log(value))
```

###Plotting for the shape versus size LMM study

estimated volume is recalculated as a Solid semi-ellipsoid around z-axis (without lobe)

volume: (2*pi*a*b*c)/3

a, b = TABW/2, PABW/2, the principal semi-axes of the base ellipse
c = BL,the principal z-semi-axe from the center of base ellipse

**2026.8 updates: using body length as estimation of size**

```{r}
df_lmw <- df_lm %>%
  dplyr::select(-c(plane, log_value)) %>%
  group_by(sample_id, ind) %>%
  pivot_wider(
    names_from = measure_name,
    values_from = value
  ) 

# previous version: using estimated volume as estimation of size
#df_lmw <- df_lmw %>% 
#  mutate(vol = (2 *(TABW/2) * (PABW/2)*BL*pi )/3)
```

#### With all variables:

```{r}
# standardize the body size through body length
measure_cols <- colnames(df_lmw[,7:23])
df_iso <-  df_lmw  %>%  
  mutate(across(all_of(measure_cols), 
                ~ .x/BL, 
                .names = "sd_{.col}"))
```

add stage
```{r}
df_pt <- read.csv("/Volumes/Yinghui/202603-dev-seq-analysis/data/202509-10-dev-tentacle.csv",header = TRUE)

df_pt <- df_pt %>% dplyr::select(c("sample_id","exp_day","tentacle"))

df_iso <- left_join(df_iso, df_pt, by = c("sample_id","exp_day"))

```

```{r}
df_iso <- df_iso %>%
mutate(stage = case_when(
is.na(ALO) & is.na(LL)  ~ "cydippid",
is.na(ALO) & !is.na(LL) ~ "early_transition",
!is.na(ALO) & !is.na(LL) & tentacle == TRUE ~ "late_transition",
!is.na(ALO) & !is.na(LL) & tentacle == FALSE ~ "lobate",
TRUE                     ~ "unclassified"
))
```

```{r}
saveRDS(df_lm, file = "/Volumes/Yinghui/202603-dev-seq-analysis/data/df_lm.rds" )
saveRDS(df_lmw, file = "/Volumes/Yinghui/202603-dev-seq-analysis/data/df_lmw.rds" )
saveRDS(df_iso, file = "/Volumes/Yinghui/202603-dev-seq-analysis/data/df_iso.rds")
```

