---
title: "202512_morph_lm"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "D:/202512-dev-seq-analyses/")

source("data_import.R")
source("PA_measurement.R")
source("TA_measurement.R")
source("OA_measurement.R")
```

```{r}
library("geomorph")
library("dplyr")
options(rgl.useNULL=TRUE)
library("rgl")
library("tidyverse")
library("gtools")
library(lubridate)
library(gridExtra)
library(factoextra)
library(ggplot2)
library(clustvarsel)
library(mclust)
library(ggpubr)
library(svglite)
```


# Import data file

Dir:"/Volumes/Yinghui/202512-dev-seq-analyses"

```{r}
# df <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds")
#df <- readRDS("D:/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds")

files <- c("D:/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds", "D:/202603-dev-seq-analysis/data/202509-dev-seq-v6.rds")
df <- lapply(files, readRDS) %>% bind_rows()
```



```{r}
# convert the lm to matrix for each file.

specs_list <- df %>%
  group_by(temp, biorep, ind, plane, file) %>%
  group_map(~ {
    id <- .y$file
    # Create a sub-list for this specific identifier
    spec <- list(
      matrix = as.matrix(.x[, c("X", "Y")]),
      metadata = as.list(.y)
    )
    
    setNames(list(spec), id)
  }) %>%
  flatten()

```

```{r}
# add lm 13 in the TA plane
specs_list<-add_lm13(specs_list)
```


# Plot setting
```{r}
palette1 <- c("#343434","#A3CEF1","#3880ba","#224f73","#000000","#CCCAC2","#837D72","#ede7da")
```

#linear measurements
```{r}
measurement_accumulator <- list()

for (id in names(specs_list)) {
  
  # 1. Extract components correctly
  spec   <- specs_list[[id]]
  mat    <- spec$matrix
  plane  <- spec$metadata$plane
  ind    <- spec$metadata$ind
  temp <- spec$metadata$temp
  biorep <- spec$metadata$biorep
  file <- spec$metadata$file
  measurer <- spec$metadata$measurer
  
 # print(file)
  # 2. Initialize a placeholder for this specific iteration's results
  res <- tibble()
  
 # print(file)
  # 3. Apply custom "Design" logic
  if (plane == "PA") {
    res <- tibble(
      measure_name = c("ALO", "AW", "WAE", "ATCL", "PABW"),
      value = c(ALO_fun(mat), AW_fun(mat), WAE_fun(mat), ATCL_fun(mat), PABW_fun(mat))
    )
    
  } else if (plane == "TA") {
    res <- tibble(
      measure_name = c("PL", "BL", "TCL", "LTM", "LL", "TABW", "AECL","TATB"),
      value = c(PL_fun(mat), BL_fun(mat), TCL_fun(mat), LTM_fun(mat), 
                LL_fun(mat), TABW_fun(mat), AECL_fun(mat), TATB_fun(mat))
    )
    
  } else if (plane == "OA") {
    res <- tibble(
      measure_name = c("TBL", "PW", "ATAC", "AEAC"),
      value = c(TBL_fun(mat), PW_fun(mat), ATAC_fun(mat), AEAC_fun(mat))
    )
  }
  
  # print(file)
  # print(row(mat))
  # print(file)
  # 4. Add metadata and store in accumulator
  # Only add if we actually generated results
  if (nrow(res) > 0) {
    measurement_accumulator[[id]] <- res %>%
      mutate(
        measurer = measurer,
        sample_id = str_extract(basename(file), "^\\d+"),
        ind = ind,
        plane = plane,
        temp = temp,
        biorep = biorep
      )
  }
}
# 5. Bind into one large dataframe
df_lm <- bind_rows(measurement_accumulator)

df_lm <- df_lm %>% mutate(
    # Convert from microns (or current unit) to mm
    value = value / 1000
  ) 
```

Removing the value of lobate and auricle when those structures are not present
standard:
ALO 0.07mm
AW 0.07mm
LTM 0.07mm
LL 0.07mm
```{r}
threshold <- 0.07

df_lm <- df_lm %>% mutate(across(
    c(value), 
    ~ if_else(measure_name %in% c("ALO", "AW", "LTM", "LL") & .x < threshold, NA_real_, .x)
  ))
  
```

convert the sample_id to date
```{r}
# Your sample data
date_strings <- df_lm$sample_id

# 1. Convert to Date objects (assuming the year 2026)
# We use parse_date_time to handle the MMDD format
dates <- parse_date_time(paste0(date_strings, "2026"), orders = "mdy")

# 2. Define your starting date (the first date in your list)
start_date <- dates[1]

# 3. Calculate days elapsed
days_diff <- as.numeric(difftime(dates, start_date, units = "days"))

# Results
df <- data.frame(
  Original = date_strings,
  Formatted_Date = dates,
  Days_Since_Start = days_diff
)

df_lm$exp_day <- df$Days_Since_Start
```
change the sample id to individual
```{r}
df_lm <- df_lm  %>%
  mutate(sample_id = paste(ind, temp, biorep, sep = "-"))
```

## allometric dev visualisation

### obtain the natural log of the value
Common practice when studying allometric development because it is easier to deal with proportional variation.

```{r}
#create a new column with log value
df_lm <- df_lm %>% mutate(log_value = log(value))
```



### plot allometric dev

Using body length as the x axis
only focusing on the 20C group
```{r}
df_lmw <- df_lm %>%
  dplyr::select(-c(plane, log_value)) %>%
  group_by(sample_id, ind) %>%
  pivot_wider(
    names_from = measure_name,
    values_from = value
  ) 

```
plot by individual

## plot all measurements together on body length
```{r}
p1 <- ggplot(df_lmw, aes(x = log(BL), y = log(TABW))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p1

p2 <- ggplot(df_lmw, aes(x = BL, y = AW)) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point(aes(color = sample_id))+
  theme(legend.position = "none")
p2

p3 <- ggplot(df_lmw, aes(x = log(BL), y = log(PABW))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p3

p4 <- ggplot(df_lmw, aes(x = log(BL), y = log(AEAC))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p4

p5 <- ggplot(df_lmw, aes(x = log(BL), y = log(AECL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p5

p6 <- ggplot(df_lmw, aes(x = log(BL), y = log(ATAC))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p6
p7 <- ggplot(df_lmw, aes(x = log(BL), y = log(ATCL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p7

p8 <- ggplot(df_lmw, aes(x = log(BL), y = log(LL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p8

p9 <- ggplot(df_lmw, aes(x = log(BL), y = log(LTM))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p9

p10 <- ggplot(df_lmw, aes(x = log(BL), y = log(PL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p10

p11 <- ggplot(df_lmw, aes(x = log(BL), y = log(PW))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p11

p12 <- ggplot(df_lmw, aes(x = log(BL), y = log(TATB))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p12

p13 <- ggplot(df_lmw, aes(x = log(BL), y = log(TBL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p13

p14 <- ggplot(df_lmw, aes(x = log(BL), y = log(TCL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p14

```

## plot all measurements together on body width
```{r}
p1 <- ggplot(df_lmw, aes(x = log(TABW), y = log(BL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p1

p2 <- ggplot(df_lmw, aes(x = log(TABW), y = log(AW))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p2

p3 <- ggplot(df_lmw, aes(x = log(TABW), y = log(PABW))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p3

p4 <- ggplot(df_lmw, aes(x = log(TABW), y = log(AEAC))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p4

p5 <- ggplot(df_lmw, aes(x = log(TABW), y = log(AECL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p5

p6 <- ggplot(df_lmw, aes(x = log(TABW), y = log(ATAC))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p6
p7 <- ggplot(df_lmw, aes(x = log(TABW), y = log(ATCL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p7

p8 <- ggplot(df_lmw, aes(x = log(TABW), y = log(LL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p8

p9 <- ggplot(df_lmw, aes(x = log(TABW), y = log(LTM))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p9

p10 <- ggplot(df_lmw, aes(x = log(TABW), y = log(PL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p10

p11 <- ggplot(df_lmw, aes(x = log(TABW), y = log(PW))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p11

p12 <- ggplot(df_lmw, aes(x = log(TABW), y = log(TATB))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p12

p13 <- ggplot(df_lmw, aes(x = log(TABW), y = log(TBL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p13

p14 <- ggplot(df_lmw, aes(x = log(TABW), y = log(TCL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p14

```
## plot all measurements together on centroid

```{r}
df_lmw <- df_lmw %>% 
  mutate(centroid = (2 *(TABW/2) * (PABW/2)*BL )/3)


measure_cols <- colnames(df_lmw[,6:22])
df_iso <-  df_lmw  %>%  
  mutate(across(all_of(measure_cols), 
                ~ log(.x/centroid), 
                .names = "shape_{.col}"))
```

```{r}
p1 <- ggplot(df_iso, aes(x = log(centroid), y = log(BL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p1

p2 <- ggplot(df_iso, aes(x = log(centroid), y = log(AW))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p2

p3 <- ggplot(df_iso, aes(x = log(centroid), y = log(PABW))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p3

p4 <- ggplot(df_lmw, aes(x = log(centroid), y = log(AEAC))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p4

p5 <- ggplot(df_lmw, aes(x = log(centroid), y = log(AECL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p5

p6 <- ggplot(df_lmw, aes(x = log(centroid), y = log(ATAC))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p6
p7 <- ggplot(df_lmw, aes(x = log(centroid), y = log(ATCL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p7

p8 <- ggplot(df_lmw, aes(x = log(centroid), y = log(LL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p8

p9 <- ggplot(df_lmw, aes(x = log(centroid), y = log(LTM))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p9

p10 <- ggplot(df_lmw, aes(x = log(centroid), y = log(PL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p10

p11 <- ggplot(df_lmw, aes(x = log(centroid), y = log(PW))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p11

p12 <- ggplot(df_lmw, aes(x = log(centroid), y = log(TATB))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p12

p13 <- ggplot(df_lmw, aes(x = log(centroid), y = log(TBL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p13

p14 <- ggplot(df_lmw, aes(x = log(centroid), y = log(TCL))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p14

p15 <- ggplot(df_iso, aes(x = log(centroid), y = log(TABW))) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p15

```

```{r}
p1 <- ggplot(df_iso, aes(x = log(centroid), y = shape_BL)) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p1

p2 <- ggplot(df_iso, aes(x = log(centroid), y = shape_AW)) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p2

p3 <- ggplot(df_iso, aes(x = log(centroid), y = shape_PABW)) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p3

p4 <- ggplot(df_iso, aes(x = log(centroid), y = shape_AEAC)) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p4

p5 <- ggplot(df_iso, aes(x = log(centroid), y = shape_AECL)) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p5

p6 <- ggplot(df_iso, aes(x = log(centroid), y = shape_ATAC)) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p6
p7 <- ggplot(df_iso, aes(x = log(centroid), y = shape_ATCL)) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p7

p8 <- ggplot(df_iso, aes(x = log(centroid), y = shape_LL)) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p8

p9 <- ggplot(df_iso, aes(x = log(centroid), y = shape_LTM)) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p9

p10 <- ggplot(df_iso, aes(x = log(centroid), y = shape_PL)) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p10

p11 <- ggplot(df_iso, aes(x = log(centroid), y = shape_PW)) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p11

p12 <- ggplot(df_iso, aes(x = log(centroid), y = shape_TATB)) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p12

p13 <- ggplot(df_iso, aes(x = log(centroid), y = shape_TBL)) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p13

p14 <- ggplot(df_iso, aes(x = log(centroid), y = shape_TCL)) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p14

p15 <- ggplot(df_iso, aes(x = log(centroid), y = shape_TABW)) +
  geom_smooth(method = "lm", linewidth = 1) +
  geom_point()
p15

```