---
title: "202512_morph_lm"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("data_import.R")
source("PA_measurement.R")
source("TA_measurement.R")
source("OA_measurement.R")
```

```{r}
library("geomorph")
library("dplyr")
options(rgl.useNULL=TRUE)
library("rgl")
library("tidyverse")
library("gtools")
library(lubridate)
library(gridExtra)
library(factoextra)
library(ggplot2)
library(clustvarsel)
library(mclust)
library(ggpubr)
library(svglite)
```


# Import data file

**need to set working dir here**
```{r}
df <- readRDS("/Volumes/Yinghui/202512-dev-seq-analyses/data/20C-202509-dev-seq-v5.rds")
```



```{r}
# convert the lm to matrix for each file.

specs_list <- df %>%
  group_by(temp, biorep,ind, plane, file) %>%
  group_map(~ {
    id <- paste(.y$temp,.y$biorep, .y$ind, .y$plane, .y$file, sep = "_")
    
    # Create a sub-list for this specific identifier
    spec <- list(
      matrix = as.matrix(.x[, c("X", "Y")]),
      metadata = as.list(.y)
    )
    
    setNames(list(spec), id)
  }) %>%
  flatten()

```

Temporary issue

```{r}
# 1. missing 1 landmark

t <- c(4607.161391, 2099.636741)
specs_list[["20_202509-dev-seq_11_TA_1007-TA-2.csv"]][["matrix"]] <- rbind(specs_list[["20_202509-dev-seq_11_TA_1007-TA-2.csv"]][["matrix"]], t)

rownames(specs_list[["20_202509-dev-seq_11_TA_1007-TA-2.csv"]][["matrix"]]) <- NULL

# 2. 1 csv has . instead of - between plane and date
specs_list[["20_202509-dev-seq_2_OA_1011.OA-1X.csv"]][["metadata"]][["file"]] <- "1011-OA-1X.csv"
```

```{r}
# add lm 10 in the TA plane
specs_list<-add_lm10(specs_list)
```

# Plot setting
```{r}
palette1 <- c("#343434","#A3CEF1","#3880ba","#224f73","#000000","#CCCAC2","#837D72","#ede7da")
```

#linear measurements
```{r}
measurement_accumulator <- list()

for (id in names(specs_list)) {
  
  # 1. Extract components correctly
  spec   <- specs_list[[id]]
  mat    <- spec$matrix
  plane  <- spec$metadata$plane
  ind    <- spec$metadata$ind
  temp <- spec$metadata$temp
  biorep <- spec$metadata$biorep
  file <- spec$metadata$file
  
  # 2. Initialize a placeholder for this specific iteration's results
  res <- tibble()

  # 3. Apply custom "Design" logic
  if (plane == "PA") {
    res <- tibble(
      measure_name = c("ALO", "AW", "WAE", "ATCL"),
      value = c(ALO_fun(mat), AW_fun(mat), WAE_fun(mat), ATCL_fun(mat))
    )
    
  } else if (plane == "TA") {
    res <- tibble(
      measure_name = c("PL", "BL", "TCL", "LTM", "LL", "TABW"),
      value = c(PL_fun(mat), BL_fun(mat), TCL_fun(mat), LTM_fun(mat), LL_fun(mat), TABW_fun(mat))
    )
    
  } else if (plane == "OA") {
    res <- tibble(
      measure_name = c("TBL", "PW", "ATAC", "AEAC"),
      value = c(TBL_fun(mat), PW_fun(mat), ATAC_fun(mat), AEAC_fun(mat))
    )
  }

  # 4. Add metadata and store in accumulator
  # Only add if we actually generated results
  if (nrow(res) > 0) {
    measurement_accumulator[[id]] <- res %>%
      mutate(
        sample_id = str_remove(file, "-.*"),
        ind = ind,
        plane = plane,
        temp = temp,
        biorep = biorep
      )
  }
}

# 5. Bind into one large dataframe
df_lm <- bind_rows(measurement_accumulator)

df_lm <- df_lm %>% mutate(
    # Convert from microns (or current unit) to mm
    value = value / 1000
  ) 
```

Removing the value of lobate and auricle when those structures are not present
```{r}
df_onlyLA <- df_lm %>% group_by(measure_name) %>% filter(measure_name %in% c("LL", "LTM")) %>% ungroup()

ggplot(df_onlyLA,aes(x=sample_id, y = value, color = measure_name))+
  geom_point(size = 1) +
  theme_classic() + 
  scale_fill_manual(values = palette1)
```
set the value before 0915 as NA
get rig of small value in the initiation of lobe anauricle development
```{r}
df_lm <- df_lm  %>% 
  mutate(
    value = case_when(
      sample_id %in% c("0905", "0907", "0909", "0911", "0913","0915") & measure_name %in% c("LL", "LTM","ALO", "AW") ~ NA,
      sample_id %in% c( "0917", "0919") &  measure_name %in% c("ALO", "AW") ~ NA,
      TRUE ~ value
    )
  ) 

#df_lm <- drop_na(df_lm)
```

## allometric dev visualisation

### obtain the natural log of the value
Common practice when studying allometric development because it is easier to deal with proportional variation.

```{r}
#create a new column with log value
df_lm <- df_lm %>% mutate(log_value = log(value))
```


### plot allometric dev

Using body length as the x axis
```{r}
df_BL <- df_lm %>%
  # 1. Create a lookup of just the BL values
  left_join(
    df_lm %>% 
      filter(measure_name == "BL") %>% 
      select(sample_id, ind, bl_value = value, bl_log = log_value), 
    by = c("sample_id","ind")
  ) %>%
  filter(!(measure_name %in% c("BL")))
 # filter(!(measure_name %in% c("BL","PL","TABW")))

```

```{r}
p <- ggplot(df_BL, aes(x = bl_log, y = log_value, color = measure_name)) +
  geom_point(size = 1) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  theme_classic() +
   scale_color_viridis_d()+ scale_fill_hue(c = 70)

p
ggsave("/Volumes/Yinghui/202512-dev-seq-analyses/fig/dev.svg", width=10, height=8)
```

```{r}
p <- p + stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))
p
```
```{r}
measurement <- c("AEAC","ATAC")
oa <- df_BL %>%
  filter(measure_name %in% measurement) %>% # Only keeps these three
  ggplot(aes(x = bl_log, y = log_value, color = measure_name)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_classic() +scale_color_manual(values = c("AEAC" = "#787A7D", "ATAC" = "#393D3F"))+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))
oa
```

```{r}
measurement <- c("TABW","TCL","WAE","PW")
fig_total <- df_BL %>%
  filter(measure_name %in% measurement) %>% # Only keeps these three
  ggplot(aes(x = bl_log, y = log_value, color = measure_name)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_classic() +scale_color_manual(values = c("TABW" = "#546A7B", "TCL" = "#8FABB7", "WAE" = "#384F57", "PW" = "#8DB2F9"))+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))

```

```{r}
measurement <- c("LL","LTM")
fig_lobe <- df_BL %>%
  filter(measure_name %in% measurement) %>% # Only keeps these three
  ggplot(aes(x = bl_log, y = log_value, color = measure_name)) +
  geom_point() +
  geom_smooth(method="lm") +
  theme_classic() +scale_color_manual(values = c("LL" = "#627F7F", "LTM" = "#94A29C"))+
stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))

fig_lobe
```
```{r}
measurement <- c("ALO","AW")
fig_auricle <- df_BL %>%
  filter(measure_name %in% measurement) %>% # Only keeps these three
  ggplot(aes(x = bl_log, y = log_value, color = measure_name)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic() +scale_color_manual(values = c("ALO" = "#C6C5B9", "AW" = "#A29F90")) + stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))

#fig_total
fig_auricle
```


combine plots
```{r}
dev<-ggarrange(fig_total, oa, fig_auricle, fig_lobe, ncol = 2, nrow = 2)
ggsave(file="/Volumes/Yinghui/202512-dev-seq-analyses/fig/dev.svg", plot=dev)
```
## Standardization and then study allometric development


### Removing isometry through Calculate centroid through geometric mean
isometric component of shape (i.e., the shape that changes in a 1:1 proportion with size) is removed

To approximate this effect in the LMM context, we used an approach that is analogous to centroid size extraction by deriving the geometric mean of all variables of a specimen as the equivalent of that specimen's centroid size and using log-shape ratios (log10[measurement/geometric mean]) as isometry-free shape variables. (Viacava 2023)

```{r}
df_lmw_nolog <- df_lm %>%
  select(-c(plane, log_value)) %>%
  group_by(sample_id, ind) %>%
  pivot_wider(
    names_from = measure_name,
    values_from = value
  )
```

```{r}
df_lmw <- df_lm %>%
  select(-c(plane, value)) %>%
  group_by(sample_id, ind) %>%
  pivot_wider(
    names_from = measure_name,
    values_from = log_value
  )
```

Starting from the day that both auricle and lobe start to present
```{r}
df_LA <- df_lmw_nolog[,5:18]

measure_cols <- colnames(df_LA)

df_lmw_nolog <- drop_na(df_lmw_nolog)

df_iso <- df_lmw_nolog %>% 
  ungroup() %>%
  mutate(centroid = exp(rowMeans(log(select(., TBL:TABW))))) 

  df_iso <-  df_iso  %>%  ungroup() %>%
  mutate(across(all_of(measure_cols), 
                ~ log(.x/centroid,10), 
                .names = "shape_{.col}"))
```

```{r}
# Create a matrix of your shape variables
shape_matrix <- df_iso %>%
  select(starts_with("shape_")) %>%
  as.matrix()

# Add the centroid back to a clean modeling data frame
model_data <- data.frame(
  centroid = df_iso$centroid,
  shape_matrix,
  ind = df_iso$ind
)

library(RRPP)
# Fit the multivariate linear model
# We typically log-transform the centroid for allometric studies
fit <- lm.rrpp(shape_matrix ~ log(centroid)*ind, SS.type = "I", 
data = model_data, print.progress = FALSE, iter = 999)
summary(fit)
anova(fit)

cor(model_data[,1:15])
```

#### With all variables:

```{r}
df_LA <- df_lmw[,5:18]

measure_cols <- colnames(df_LA)

df_iso <- df_lmw %>% 
  ungroup() %>%
  mutate(centroid = rowMeans(select(., TBL:TABW), na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(across(all_of(measure_cols), 
                ~ .x - centroid, 
                .names = "shape_{.col}"))
```


```{r}
model_LA = glm(formula = centroid ~ shape_TBL + shape_PW + shape_ATAC +shape_AEAC + shape_ALO+ shape_AW+ shape_WAE + shape_ATCL + shape_PL+ shape_BL+ shape_TCL + shape_LTM + shape_LL + shape_TABW, data = df_iso)
plot(model_LA)
summary(model_LA)
```



to assess the effect of allometry on shape variation:

For LMM, we regressed the linear data against the natural logarithm of the geometric mean with the lm.rrpp function of the “RRPP” package (version 1.3.1) (Collyer & Adams, 2018). We considered both centroid size and geometric mean as proxies for size in the context of geometric and LMM. (Viacava 2023)


## Clustering
### GMM analysis

Select variables that are most significant to the model

First we look at the result obtained using the function Mclust from the mclust package,
with the best model selected by BIC for clustering on all the variables, allowing all possible
parameterizations and the number of groups

```{r}
# pivot wide the data, so that each measurement is in a separate column
df_lmw <- df_lm %>%
  select(-c(plane, value)) %>%
  group_by(sample_id, ind) %>%
  pivot_wider(
    names_from = measure_name,
    values_from = log_value
  )
```

#### With all variables:

```{r}
df_LA <- df_lmw[,5:18]
mod1 <- Mclust(df_LA,G=1:8)
summary(mod1)
```

```{r}
out1 <- clustvarsel(df_LA, G=4)
out1
```
Selected subset: ATAC, AW, ALO, LL, LTM, TCL, ATCL

```{r}
mod1_2 <-Mclust(df_LA[, out1$subset],G=4)
summary(mod1_2)
```

```{r}
fviz_mclust(mod1_2,"classification",  geom = "point", palette = palette1)
plot(mod1_2, what = "classification", colors = palette1)
```

```{r}
palette1 <- sample(palette1)
p1 <- fviz_mclust(mod1_2,"classification",  geom = "point", palette = palette1)

#svglite("/Volumes/Yinghui/2202512-dev-seq-analyses/fig/class1.svg", width = 10, height = 10)
plot(mod1_2, what = "classification", colors = palette1)
dev.off()

```

```{r}
ggsave(file="/Volumes/Yinghui/202512-dev-seq-analyses/fig/GMM1.svg", plot=p1, width=10, height=10)
```

##### with iso (shape only)
```{r}
mod1_1 <- Mclust(df_iso[,19:33])
summary(mod1_1)
```

#### With no auricle and lobe related measurements:
```{r}
df_noLA <- df_lmw[,!names(df_lmw) %in% c("ALO", "AW", "LL", "LTM", "sample_id", "temp", "ind", "biorep")]

#examine their correlation
m <- cor(df_noLA)
# Convert matrix to data frame and reshape
cor_df <- as.data.frame(m) 

a <- which(cor_df>0.99, arr.ind=T)
row.names(cor_df[a[,1],] )
names(cor_df[a[,2]])  
```

```{r}
mod2 <- Mclust(df_noLA)
summary(mod2)
```




Selecting variable useful for clustering
```{r}
out2 <- clustvarsel(df_noLA)
out2
```


Construct the clustering model again
```{r}
mod2_2 <-Mclust(df_noLA[, out2$subset])
summary(mod2_2)
```

```{r}
palette1 <- sample(palette1)
p3 <- fviz_mclust(mod2_2,"classification",  geom = "point", palette = palette1)

svglite("/Volumes/Yinghui/202512-dev-seq-analyses/fig/class2.svg", width = 10, height = 10)
plot(mod2_2, what = "classification", colors = palette1)
dev.off()

```

```{r}
ggsave(file="/Volumes/Yinghui/202512-dev-seq-analyses/fig/GMM2.svg", plot=p3, width=10, height=10)
```


### K-means
for group w.o lobe or auricle
```{r}
km.res <- kmeans(df_noLA,5)
print(km.res)
```
### PCA

```{r}
PCA2 <- prcomp(select_if(df_noLA, is.numeric), scale. = FALSE)
summary(PCA2)
fviz_pca_ind(PCA2)
```
