---
title: "202512_morph_lm"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "D:/202512-dev-seq-analyses/")

source("data_import.R")
source("PA_measurement.R")
source("TA_measurement.R")
source("OA_measurement.R")
```

```{r}
library("geomorph")
library("dplyr")
options(rgl.useNULL=TRUE)
library("rgl")
library("tidyverse")
library("gtools")
library(lubridate)
library(gridExtra)
library(factoextra)
library(ggplot2)
library(clustvarsel)
library(mclust)
library(ggpubr)
library(svglite)
```


# Import data file

Dir:"/Volumes/Yinghui/202512-dev-seq-analyses"

```{r}
df <- readRDS("/Volumes/Yinghui/202512-dev-seq-analyses/data/20C-202509-dev-seq-v5.rds")
```



```{r}
# convert the lm to matrix for each file.

specs_list <- df %>%
  group_by(temp, biorep,ind, plane, file) %>%
  group_map(~ {
    id <- paste(.y$temp,.y$biorep, .y$ind, .y$plane, .y$file, sep = "_")
    
    # Create a sub-list for this specific identifier
    spec <- list(
      matrix = as.matrix(.x[, c("X", "Y")]),
      metadata = as.list(.y)
    )
    
    setNames(list(spec), id)
  }) %>%
  flatten()

```

Temporary issue

```{r}
# 1. missing 1 landmark

t <- c(4607.161391, 2099.636741)
specs_list[["20_202509-dev-seq_11_TA_1007-TA-2.csv"]][["matrix"]] <- rbind(specs_list[["20_202509-dev-seq_11_TA_1007-TA-2.csv"]][["matrix"]], t)

rownames(specs_list[["20_202509-dev-seq_11_TA_1007-TA-2.csv"]][["matrix"]]) <- NULL

# 2. 1 csv has . instead of - between plane and date
specs_list[["20_202509-dev-seq_2_OA_1011.OA-1X.csv"]][["metadata"]][["file"]] <- "1011-OA-1X.csv"
```

```{r}
# add lm 10 in the TA plane
specs_list<-add_lm10(specs_list)
```

# Plot setting
```{r}
palette1 <- c("#343434","#A3CEF1","#3880ba","#224f73","#000000","#CCCAC2","#837D72","#ede7da")
```

#linear measurements
```{r}
measurement_accumulator <- list()

for (id in names(specs_list)) {
  
  # 1. Extract components correctly
  spec   <- specs_list[[id]]
  mat    <- spec$matrix
  plane  <- spec$metadata$plane
  ind    <- spec$metadata$ind
  temp <- spec$metadata$temp
  biorep <- spec$metadata$biorep
  file <- spec$metadata$file
  
  # 2. Initialize a placeholder for this specific iteration's results
  res <- tibble()

  # 3. Apply custom "Design" logic
  if (plane == "PA") {
    res <- tibble(
      measure_name = c("ALO", "AW", "WAE", "ATCL"),
      value = c(ALO_fun(mat), AW_fun(mat), WAE_fun(mat), ATCL_fun(mat))
    )
    
  } else if (plane == "TA") {
    res <- tibble(
      measure_name = c("PL", "BL", "TCL", "LTM", "LL", "TABW"),
      value = c(PL_fun(mat), BL_fun(mat), TCL_fun(mat), LTM_fun(mat), LL_fun(mat), TABW_fun(mat))
    )
    
  } else if (plane == "OA") {
    res <- tibble(
      measure_name = c("TBL", "PW", "ATAC", "AEAC"),
      value = c(TBL_fun(mat), PW_fun(mat), ATAC_fun(mat), AEAC_fun(mat))
    )
  }

  # 4. Add metadata and store in accumulator
  # Only add if we actually generated results
  if (nrow(res) > 0) {
    measurement_accumulator[[id]] <- res %>%
      mutate(
        sample_id = str_remove(file, "-.*"),
        ind = ind,
        plane = plane,
        temp = temp,
        biorep = biorep
      )
  }
}

# 5. Bind into one large dataframe
df_lm <- bind_rows(measurement_accumulator)

df_lm <- df_lm %>% mutate(
    # Convert from microns (or current unit) to mm
    value = value / 1000
  ) 
```

Removing the value of lobate and auricle when those structures are not present
```{r}
df_onlyLA <- df_lm %>% group_by(measure_name) %>% filter(measure_name %in% c("LL", "LTM")) %>% ungroup()

ggplot(df_onlyLA,aes(x=sample_id, y = value, color = measure_name))+
  geom_point(size = 1) +
  theme_classic() + 
  scale_fill_manual(values = palette1)
```
set the value before 0915 as NA
get rig of small value in the initiation of lobe anauricle development
```{r}
df_lm <- df_lm  %>% 
  mutate(
    value = case_when(
      sample_id %in% c("0905", "0907", "0909", "0911", "0913","0915") & measure_name %in% c("LL", "LTM","ALO", "AW") ~ NA,
      sample_id %in% c( "0917", "0919") &  measure_name %in% c("ALO", "AW") ~ NA,
      TRUE ~ value
    )
  ) 

#df_lm <- drop_na(df_lm)
```

## allometric dev visualisation

### obtain the natural log of the value
Common practice when studying allometric development because it is easier to deal with proportional variation.

```{r}
#create a new column with log value
df_lm <- df_lm %>% mutate(log_value = log(value))
```


### plot allometric dev

Using body length as the x axis
```{r}
df_BL <- df_lm %>%
  # 1. Create a lookup of just the BL values
  left_join(
    df_lm %>% 
      filter(measure_name == "BL") %>% 
      select(sample_id, ind, bl_value = value, bl_log = log_value), 
    by = c("sample_id","ind")
  ) %>%
  filter(!(measure_name %in% c("BL")))
 # filter(!(measure_name %in% c("BL","PL","TABW")))

```

plot all measurements together
```{r}
p <- ggplot(df_BL, aes(x = bl_log, y = log_value, color = measure_name)) +
  geom_point(size = 1) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  theme_classic() +
   scale_color_viridis_d()+ scale_fill_hue(c = 70)

p
ggsave("/Volumes/Yinghui/202512-dev-seq-analyses/fig/dev.svg", width=10, height=8)
```

Add equation on the linear regression line on the plot
```{r}
p <- p + stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))
p
```

Group the measurements so that the figure will be clearer.
measurements on the aboral plane and related to the canals length
```{r}
measurement <- c("AEAC","ATAC")
oa <- df_BL %>%
  filter(measure_name %in% measurement) %>% # Only keeps these three
  ggplot(aes(x = bl_log, y = log_value, color = measure_name)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_classic() +scale_color_manual(values = c("AEAC" = "#787A7D", "ATAC" = "#393D3F"))+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))
oa
```

measurements related to the most obvious sturctures
```{r}
measurement <- c("TABW","TCL","WAE","PW")
fig_total <- df_BL %>%
  filter(measure_name %in% measurement) %>% # Only keeps these three
  ggplot(aes(x = bl_log, y = log_value, color = measure_name)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_classic() +scale_color_manual(values = c("TABW" = "#546A7B", "TCL" = "#8FABB7", "WAE" = "#384F57", "PW" = "#8DB2F9"))+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))

```

measurements related to the lobes
```{r}
measurement <- c("LL","LTM")
fig_lobe <- df_BL %>%
  filter(measure_name %in% measurement) %>% # Only keeps these three
  ggplot(aes(x = bl_log, y = log_value, color = measure_name)) +
  geom_point() +
  geom_smooth(method="lm") +
  theme_classic() +scale_color_manual(values = c("LL" = "#627F7F", "LTM" = "#94A29C"))+
stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))

fig_lobe
```

measurements related to the auricles
```{r}
measurement <- c("ALO","AW")
fig_auricle <- df_BL %>%
  filter(measure_name %in% measurement) %>% # Only keeps these three
  ggplot(aes(x = bl_log, y = log_value, color = measure_name)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic() +scale_color_manual(values = c("ALO" = "#C6C5B9", "AW" = "#A29F90")) + stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))

#fig_total
fig_auricle
```


combine plots
```{r}
dev<-ggarrange(fig_total, oa, fig_auricle, fig_lobe, ncol = 2, nrow = 2)
ggsave(file="/Volumes/Yinghui/202512-dev-seq-analyses/fig/dev.svg", plot=dev)
```

creating a similar figure but with width as x-axis (because in the temperature experiment it will be the diameter)

```{r}
#Using body width as the x axis

df_TABW <- df_lm %>%
  # 1. Create a lookup of just the BL values
  left_join(
    df_lm %>% 
      filter(measure_name == "TABW") %>% 
      select(sample_id, ind, bw_value = value, bw_log = log_value), 
    by = c("sample_id","ind")
  ) %>%
  filter(!(measure_name %in% c("TABW")))

# for canals in OA plane
measurement <- c("AEAC","ATAC")
oa <- df_TABW %>%
  filter(measure_name %in% measurement) %>% # Only keeps these three
  ggplot(aes(x = bw_log, y = log_value, color = measure_name)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_classic() +scale_color_manual(values = c("AEAC" = "#787A7D", "ATAC" = "#393D3F"))+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))


# measurements related to the most obvious sturctures
measurement <- c("BL","TCL","WAE","PW")
fig_total <- df_TABW %>%
  filter(measure_name %in% measurement) %>% # Only keeps these three
  ggplot(aes(x = bw_log, y = log_value, color = measure_name)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_classic() +scale_color_manual(values = c("BL" = "#546A7B", "TCL" = "#8FABB7", "WAE" = "#384F57", "PW" = "#8DB2F9"))+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))

#measurements related to the lobes
measurement <- c("LL","LTM")
fig_lobe <- df_TABW %>%
  filter(measure_name %in% measurement) %>% # Only keeps these three
  ggplot(aes(x = bw_log, y = log_value, color = measure_name)) +
  geom_point() +
  geom_smooth(method="lm") +
  theme_classic() +scale_color_manual(values = c("LL" = "#627F7F", "LTM" = "#94A29C"))+
stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))

#measurements related to the auricles

measurement <- c("ALO","AW")
fig_auricle <- df_TABW %>%
  filter(measure_name %in% measurement) %>% # Only keeps these three
  ggplot(aes(x = bw_log, y = log_value, color = measure_name)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic() +scale_color_manual(values = c("ALO" = "#C6C5B9", "AW" = "#A29F90")) + stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))


#combine plots
dev<-ggarrange(fig_total, oa, fig_auricle, fig_lobe, ncol = 2, nrow = 2)
ggsave(file="/Volumes/Yinghui/202512-dev-seq-analyses/fig/dev_tbws.svg", plot=dev)
```
###Comparing the slope based on body length v.s. 

Using ancova to determine whether there are significant differences  between slopes

```{r}
#library(car)
 # fit the model using the aov function
#df_lmwide <- df_lm %>%
#  select(-c(plane, value)) %>%
#  group_by(sample_id, ind) %>%
#  pivot_wider(
#    names_from = measure_name,
#    values_from = log_value
#  )

#model <- aov(cbind(TABW, LTM, ALO) ~ BL + Error(ind), data = df_lmwide)
#summary(model)
```


## Standardization and then study allometric development


### Removing isometry through Calculate centroid through geometric mean
isometric component of shape (i.e., the shape that changes in a 1:1 proportion with size) is removed

To approximate this effect in the LMM context, we used an approach that is analogous to centroid size extraction by deriving the geometric mean of all variables of a specimen as the equivalent of that specimen's centroid size and using log-shape ratios (log10[measurement/geometric mean]) as isometry-free shape variables. (Viacava 2023)

```{r}
# shape from
df_lmw_nolog <- df_lm %>%
  select(-c(plane, log_value)) %>%
  group_by(sample_id, ind) %>%
  pivot_wider(
    names_from = measure_name,
    values_from = value
  )
```

```{r}
# shape
df_lmw <- df_lm %>%
  select(-c(plane, value)) %>%
  group_by(sample_id, ind) %>%
  pivot_wider(
    names_from = measure_name,
    values_from = log_value
  )
```

Starting from the day that both auricle and lobe start to present
```{r}
df_LA <- df_lmw_nolog[,5:18]

measure_cols <- colnames(df_LA)

df_lmw_nolog <- drop_na(df_lmw_nolog)

df_iso <- df_lmw_nolog %>% 
  ungroup() %>%
  mutate(centroid = exp(rowMeans(log(select(., TBL:TABW))))) 

  df_iso <-  df_iso  %>%  ungroup() %>%
  mutate(across(all_of(measure_cols), 
                ~ log(.x/centroid,10), 
                .names = "shape_{.col}"))
```

```{r}
# Create a matrix of your shape variables
shape_matrix <- df_iso %>%
  select(starts_with("shape_")) %>%
  as.matrix()

# Add the centroid back to a clean modeling data frame
model_data <- data.frame(
  centroid = df_iso$centroid,
  shape_matrix,
  ind = df_iso$ind
)

library(RRPP)
# Fit the multivariate linear model
# We typically log-transform the centroid for allometric studies
fit <- lm.rrpp(shape_matrix ~ log(centroid)*ind, SS.type = "I", 
data = model_data, print.progress = FALSE, iter = 999)
summary(fit)
anova(fit)

```

####Plotting for the shape versus size LMM study

```{r}
df_lmw <- drop_na(df_lmw)
df_lmw$centroid <- log(df_iso$centroid)
df_long <- df_lmw %>% pivot_longer(
    cols = "TBL":"TABW",           # The range of columns you want to collapse
    names_to = "measurement",   # What to call the new column for the headers (TBL, PW, etc.)
    values_to = "value"         # What to call the new column for the numbers
  )

measurement <- c("BL","TABW")
p1 <-  df_long %>%# Only keeps these three
  ggplot(aes(x = centroid, y = value, color = measurement)) +
  geom_point() +
  geom_smooth(method = "lm", se=FALSE) +
  theme_classic() + stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))

p1
```
```{r}
p1 <-ggplot(df_lmw, aes(x = log(centroid), y = PL, colour = ind)) +
    geom_point(size = 2, alpha = 0.8) +
    # Add the dashed regression line
    geom_smooth(method = "lm", formula = y ~ x, color = "black") +
    theme_classic() +
    labs(x = "log Centroid size", y = "log of the pharynx length (mm)") +
    scale_color_manual(values = c("#FBD1A2", "#00B2CA", "#F79256", "#7DCFB6")) +
    theme(legend.position = "right")
#+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))
p4 <-ggplot(df_lmw, aes(x = log(centroid), y = TABW, colour = ind)) +
    geom_point(size = 2, alpha = 0.8) +
    # Add the dashed regression line
    geom_smooth(method = "lm", formula = y ~ x, color = "black") +
    theme_classic() +
    labs(x = "log Centroid size", y = "log of the body width (mm)") +
    scale_color_manual(values = c("#FBD1A2", "#00B2CA", "#F79256", "#7DCFB6")) +
    theme(legend.position = "right")
#+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))
p2 <- ggplot(df_lmw, aes(x = log(centroid), y = LL, colour = ind)) +
    geom_point(size = 2, alpha = 0.8) +
    # Add the dashed regression line
    geom_smooth(method = "lm", formula = y ~ x, color = "black") +
    theme_classic() +
    labs(x = "log Centroid size", y = "log of the lobe Width (mm)") +
    scale_color_manual(values = c("#FBD1A2", "#00B2CA", "#F79256", "#7DCFB6")) +
    theme(legend.position = "right")
#+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))
p3 <- ggplot(df_lmw, aes(x = log(centroid), y = AW, colour = ind)) +
    geom_point(size = 2, alpha = 0.8) +
    # Add the dashed regression line
    geom_smooth(method = "lm", formula = y ~ x, color = "black") +
    theme_classic() +
    labs(x = "log Centroid size", y = "log of the auricle Length (mm)") +
    scale_color_manual(values = c("#FBD1A2", "#00B2CA", "#F79256", "#7DCFB6")) +
    theme(legend.position = "right")
#+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))

p <- ggarrange(p1,p4, p2,p3, nrow = 2, ncol = 2,common.legend = TRUE)
ggsave(file="/Volumes/Yinghui/202512-dev-seq-analyses/fig/p.svg", plot=p)
```

#### With all variables:

```{r}
df_LA <- df_lmw[,5:18]

measure_cols <- colnames(df_LA)

df_iso <- df_lmw %>% 
  ungroup() %>%
  mutate(centroid = rowMeans(select(., TBL:TABW), na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(across(all_of(measure_cols), 
                ~ .x - centroid, 
                .names = "shape_{.col}"))
```


```{r}
model_LA = glm(formula = centroid ~ shape_TBL + shape_PW + shape_ATAC +shape_AEAC + shape_ALO+ shape_AW+ shape_WAE + shape_ATCL + shape_PL+ shape_BL+ shape_TCL + shape_LTM + shape_LL + shape_TABW, data = df_iso)
plot(model_LA)
summary(model_LA)
```



to assess the effect of allometry on shape variation:

For LMM, we regressed the linear data against the natural logarithm of the geometric mean with the lm.rrpp function of the “RRPP” package (version 1.3.1) (Collyer & Adams, 2018). We considered both centroid size and geometric mean as proxies for size in the context of geometric and LMM. (Viacava 2023)


## Clustering
### GMM analysis

Select variables that are most significant to the model

First we look at the result obtained using the function Mclust from the mclust package,
with the best model selected by BIC for clustering on all the variables, allowing all possible
parameterizations and the number of groups

```{r}
# pivot wide the data, so that each measurement is in a separate column

# not shape but log measurements

df_lmw <- df_lm %>%
  select(-c(plane, value)) %>%
  group_by(sample_id, ind) %>%
  pivot_wider(
    names_from = measure_name,
    values_from = log_value
  )

```

#### With all variables:

```{r}
# df_LA: with lobe and auricles
df_LA <- df_lmw[,5:18]
df_LA <- drop_na(df_LA)
mod1 <- Mclust(df_LA,G=1:8)
summary(mod1)
```

```{r}
out1 <- clustvarsel(df_LA, G=4)
out1
```
```{r}
steps_data <- as.data.frame(out1$steps.info) %>%
  mutate(
    Step = row_number(),
    # 'Variable' is usually the name of the column added/removed
    Variable = rownames(out1$steps.info)
  ) %>%
  # We want to see how much the BIC improved at each step
  rename(BIC_Change = 1)

ggplot(steps_data, aes(x = reorder(Variable, Step), y = BIC_Change)) +
  geom_segment(aes(xend = Variable, yend = 0), color = "skyblue") +
  geom_point(aes(color = BIC_Change > 0), size = 4) +
  coord_flip() +  # Flip it to read variable names easily
  theme_minimal() +
  scale_color_manual(values = c("red", "darkgreen"), labels = c("Rejection", "Selection")) +
  labs(
    title = "Clustvarsel: Variable Importance",
    subtitle = "Variables ordered by their contribution to cluster separation",
    x = "Measurement Variable",
    y = "BIC Difference (Evidence Strength)",
    color = "Decision"
  )
```
Selected subset: ATAC, AW, ALO, LL, LTM, TCL, ATCL

```{r}
mod1_2 <-Mclust(df_LA[, out1$subset])
summary(mod1_2)
```

```{r}

#fviz_mclust(mod1_2,"classification",  geom = "point", palette = palette1)
plot(mod1_2, what = "classification", colors = palette1)
plot(mod1_2, what = "BIC", colors = palette1)
```

```{r}
#palette1 <- sample(palette1)
p1 <- fviz_mclust(mod1_2,"classification",  geom = "point", palette = palette1)
p1
svglite("/Volumes/Yinghui/202512-dev-seq-analyses/fig/class1-bic.svg", width = 10, height = 10)

plot(mod1_2, what = "BIC", colors = palette1)
dev.off()

```

```{r}
ggsave(file="/Volumes/Yinghui/202512-dev-seq-analyses/fig/GMM1.svg", plot=p1, width=5, height=5)
```

```{r}
df_lmw_class <- drop_na(df_lmw)
df_lmw_class$classification <- as.factor(mod1_2$classification)


PCA2 <- prcomp(df_lmw_class[,5:18], scale. = FALSE)
summary(PCA2)
p <- fviz_pca_ind(PCA2, geom.ind = "point",
             col.ind = df_lmw_class$classification, # color by groups
             palette = c("#94A29C", "#8DB2B9", "#384F57","#546A7B"),
             addEllipses = TRUE, ellipse.type = "convex",
             legend.title = "Groups")

ggsave(file="/Volumes/Yinghui/202512-dev-seq-analyses/fig/PCA1.svg", plot=p, width=5, height=5)

```

##### with iso (shape only)
```{r}
mod1_1 <- Mclust(df_iso[,19:33])
summary(mod1_1)
```

#### With no auricle and lobe related measurements:
```{r}
# df_noLA: no lobe and auricles measurements
df_noLA <- df_lmw[,!names(df_lmw) %in% c("ALO", "AW", "LL", "LTM", "sample_id", "temp", "ind", "biorep")]

#examine their correlation
m <- cor(df_noLA)
# Convert matrix to data frame and reshape
cor_df <- as.data.frame(m) 

a <- which(cor_df>0.99, arr.ind=T)
row.names(cor_df[a[,1],] )
names(cor_df[a[,2]])  
```

```{r}
mod2 <- Mclust(df_noLA)
summary(mod2)
```




Selecting variable useful for clustering
```{r}
out2 <- clustvarsel(df_noLA)
out2
```


Construct the clustering model again
```{r}
mod2_2 <-Mclust(df_noLA[, out2$subset])
summary(mod2_2)
```

```{r}
#palette1 <- sample(palette1)
p3 <- fviz_mclust(mod2_2,"classification",  geom = "point", palette = palette1) +
  scale_shape_manual(values = c(16,0,17)) +
  scale_color_manual(values = c("#343434","#ccc9c1","#a4cdee"))
  
p3
svglite("/Volumes/Yinghui/202512-dev-seq-analyses/fig/class2.svg", width = 6, height = 6)
plot(mod2_2, what = "classification", colors = palette1)
dev.off()
plot(mod2_2, what = "classification", colors = palette1)

```

```{r}
ggsave(file="/Volumes/Yinghui/202512-dev-seq-analyses/fig/GMM2.svg", plot=p3, width=5, height=5)
```

overlap the classification back
```{r}
df_lmw$cluster <- as.factor(mod2_2$classification)

p<-ggplot(df_lmw, aes(x = BL, y = TABW, color = cluster)) +
  labs(x = "Log of body length (mm)",y = "log of body width (mm)")+
  geom_point(size = 1) +
  theme_classic() +
  scale_shape_manual(values = c(16,0,17)) +
  scale_color_manual(values = c("#343434","#ccc9c1","#a4cdee"))
ggsave(file="/Volumes/Yinghui/202512-dev-seq-analyses/fig/GM2_mapping.svg", plot=p, width=5, height=5)

```

### K-means
for group w.o lobe or auricle
```{r}
km.res <- kmeans(df_noLA,5)
print(km.res)
```
### PCA

```{r}
PCA2 <- prcomp(select_if(df_noLA, is.numeric), scale. = FALSE)
summary(PCA2)
fviz_pca_ind(PCA2)
```


