---
title: "202512_morph_lm"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}
library("geomorph")
library("dplyr")
options(rgl.useNULL=TRUE)
library("rgl")
library("tidyverse")
library("gtools")
library(lubridate)
library(gridExtra)
library(factoextra)
library(ggplot2)
library(clustvarsel)
library(mclust)
library(ggpubr)
library(svglite)
library(MASS)
```

Import the dataframe created in the load_dataframe.rmd
```{r}
df_lm <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/df_lm.rds")
df_lmw <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/df_lmw.rds")
df_iso <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/df_iso.rds")
#df_NoLA_6.4 <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/df_NoLA.rds")
#df_NoLA <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/df_NoLA.rds")
df_gmm_export <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/df_gmm_export.rds")

```


```{r}
#df_NoLA <- df_iso %>% 
#  dplyr::select(-c("ALO", "AW", "LTM", "LL", "sd_LL","sd_LTM","sd_ALO","sd_AW"))%>%
#  drop_na() %>%
#  filter(if_all(where(is.numeric),  ~ is.finite(.x) & .x != 0))
  #filter(if_all(where(is.numeric), is.finite))

#df_NoLA[,7:19]
#df_NoLA[,20:32], lm standardized by volume and logged
#df_NoLA[,39:51]
#df_gmm_export[,20:32]
df_test <-df_gmm_export[,20:32] %>% 
  mutate (stage = df_gmm_export$stage) %>% 
  dplyr::select(-c("sd_BL"))

```

# Linear discriminate analysis
```{r}
lda.stg <-lda(stage ~ ., data = df_test)
lda.stg
```

```{r}
# Extract scaling matrix and compute absolute coefficients
df_coefs <- as.data.frame(lda.stg$scaling) %>%
  tibble::rownames_to_column(var = "Feature") %>%
  mutate(
    # Create absolute values for sorting importance
    Abs_LD1 = abs(LD1)
  )

lda_con <- ggplot(df_coefs, aes(x = reorder(Feature, Abs_LD1), y = LD1, fill = LD1 > 0)) +
  geom_col(show.legend = FALSE, width = 0.6) +
  scale_fill_manual(values = palette4) +
  coord_flip() +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

lda_con
ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/lda_contribution.svg", plot=lda_con)
```

plot
```{r}
stg.sub <- 
  df_test[1:12]%>% 
  as.matrix  # cast to matrix for calculations

# calculate CV scores
CVA.scores <- stg.sub %*% lda.stg$scaling

# create data frame with scores
stg.CV <- data.frame(CVA.scores)
stg.CV$stage <- df_test$stage
```

```{r}
cva.plot <-
  ggplot(stg.CV, aes(x = LD1, y = LD2)) + 
  geom_point(aes(color=stage, shape=stage)) + 
  scale_color_manual(values = palette4)+
  labs(x = "LD1", y = "LD2") +
  coord_fixed(ratio=1) +# keep the unit scaling of the plot fixed at 1
  theme_classic()

cva.plot
```

```{r}
lda_plot <-  ggplot(stg.CV, aes(x = LD1, y = LD2)) + 
  geom_point(aes(color=stage))  + 
  scale_colour_manual(values = palette4)+
  coord_fixed(ratio=1)+
  theme_classic()# keep the unit scaling of the plot fixed at 1
lda_plot
ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/lda_plot_BL_standardized_stage.svg", plot=lda_plot)
```



```{r}
ggplot(stg.CV, aes(x = LD1)) +
  geom_density(aes(color=stage)) +
  scale_colour_manual(values = palette4)+
  labs(x = "CV1")
```

### LDA of gmm defined groups
```{r}
df_test2 <-df_gmm_export[,20:32] %>% 
  mutate (cluster = df_gmm_export$cluster_shape) %>% 
  dplyr::select(-c("sd_BL"))

lda.stg <-lda(cluster ~ ., data = df_test2)
lda.stg

stg.sub <- 
  df_test2[1:12]%>% 
  as.matrix  # cast to matrix for calculations

# calculate CV scores
CVA.scores <- stg.sub %*% lda.stg$scaling

# create data frame with scores
stg.CV <- data.frame(CVA.scores)
stg.CV$cluster <- df_test2$cluster
lda_plot <-  ggplot(stg.CV, aes(x = LD1, y = LD2)) + 
  geom_point(aes(color=cluster)) + 
  scale_colour_manual(values = palette5)+
  coord_fixed(ratio=1)+
  theme_classic()# keep the unit scaling of the plot fixed at 1
lda_plot
ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/lda_plot_BL_standardized_cluster.svg", plot=lda_plot)
```



#PCA
```{r}
#measure_cols <- colnames(df_NoLA[,7:19])
#df_NoLA <-  df_NoLA  %>%  
#  mutate(across(all_of(measure_cols), 
#                ~ log(.x/exp(mean(log(.x)))), 
#                .names = "shape_ratio_{.col}"))
#df_NoLA[,7:19]
#df_NoLA[,40:52],
#df_NoLA[,21:33]
df_test <- df_test %>% mutate(cluster = df_gmm_export$cluster_shape)
pca<- prcomp(df_test[,1:12],
                   center = TRUE,
                   scale. = TRUE)

```

## color by stage
```{r}
library(ggfortify)
pca.plot <- autoplot(pca,
                          data = df_test,
                          colour = 'stage')+
  scale_colour_manual(values = palette4)+
  theme_classic()

pca.plot
ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/pca_plot_by_stage.svg", plot=pca.plot)
```
color by cluster
```{r}
pca.plot <- autoplot(pca,
                          data = df_test,
                          colour = 'cluster')+
  scale_colour_manual(values = palette5)+
  theme_classic()

pca.plot
ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/pca_plot_by_cluster.svg", plot=pca.plot)
```

```{r}
con1 <- fviz_contrib(pca, choice = "var", axes = 1, top = 10)
con2 <- fviz_contrib(pca, choice = "var", axes = 2, top = 10)
con1
fviz_contrib(pca, choice = "var", axes = 3, top = 10)
```
identify the min and max point of PC1 and pc2
```{r}
min_idx <- which.min(pca$x[, 1])
df_gmm_export[min_idx,]

max_idx <- which.max(pca$x[, 2])
df_gmm_export[max_idx,]
```
```{r}
ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/pca_contribution1.svg", plot=con1)
ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/pca_contribution2.svg", plot=con2)
```

## PCA identify outliers
### standard way

```{r}
# remotes::install_github("privefl/bigutilsr")
library(bigutilsr)
U <- pca$x
#theme_set(bigstatsr::theme_bigstatsr(0.8))
# qplot(U[, 1], U[, 2]) + coord_equal()
# Standard outlier detection:
# criterion of being “more than 6 standard deviations away from the mean”.

outlier_idx <- unique(unlist(apply(U, 2, function(x) which(abs(x - mean(x)) > (6 * sd(x))))))
col <- rep("black", nrow(U))
col[outlier_idx] <- "red"
qplot(U[, 1], U[, 2], color = I(col), size = I(2)) + coord_equal()
```


```{r}
df_plot <-df_NoLA %>%
  mutate(
    resPC1 = pca$x[, 1])

ggplot(df_plot, aes(x = t_centroid, y = resPC1, color = stage)) +
  geom_point() +
  theme_classic() + labs(x = "size", y = "PC1 score")
```
###Robust Mahalanobis distance
```{r}

```

### using packages that based on Mahalanobis distance

```{r}
library(mdatools)
# 1. Fit mdatools PCA model and assign to 'm'
m <- mdatools::pca(df_test[1:12], scale = TRUE, ncomp = 4)

# 2. Pass 'm' to plotContributions
plotDistances(m, show.labels = TRUE)
```
## perform PCA on the residuals


```{r}
residual_data <- df_NoLA %>%
  mutate(across(all_of(measure_cols), 
    ~ residuals(lm(log(.x) ~ log(vol), na.action = na.exclude))
  )) %>%
dplyr::select(all_of(measure_cols))
```

```{r}
# Run PCA on the shape residuals
pca_residual_result <- prcomp(residual_data, center = TRUE, scale. = TRUE)

# Bind these new "Size-Free" PC scores back to your metadata
df_final_plot <- df_NoLA %>%
  mutate(
    resPC1 = pca_residual_result$x[, 1],
    resPC2 = pca_residual_result$x[, 2]
  )
```

```{r}
ggplot(df_final_plot, aes(x = vol, y = resPC1, color = stage)) +
  geom_point() +
  scale_size_continuous(range = c(2, 6)) +
  theme_classic()
```

```{r}
# Extract loadings (rotations)
loadings <- as.data.frame(pca_residual_result$rotation)
loadings$Trait <- rownames(loadings)

# Plot the structural impact of each trait on Residual PC1
ggplot(loadings, aes(x = reorder(Trait, PC1), y = PC1)) +
  geom_col(fill = "#384F57") +
  coord_flip() +
  labs(
    title = "Trait Contributions to Residual PC1",
    x = "Linear Measurement",
    y = "Loading Score (Direction of Shape Change)"
  ) +
  theme_minimal()
```
```{r}
fviz_contrib(pca_residual_result, choice = "var", axes = 1, top = 10)
fviz_contrib(pca_residual_result, choice = "var", axes = 2, top = 10)
fviz_contrib(pca_residual_result, choice = "var", axes = 3, top = 10)
```
PCA of the measurements of the residuals
```{r}
pca.plot_res <- autoplot(pca_residual_result,
                          data = df_NoLA,
                          colour = 'stage')

pca.plot_res
```
