---
title: "202512_morph_lm"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "D:/202512-dev-seq-analyses/")
```

```{r}
library("geomorph")
library("dplyr")
options(rgl.useNULL=TRUE)
library("rgl")
library("tidyverse")
library("gtools")
library(lubridate)
library(gridExtra)
library(factoextra)
library(ggplot2)
library(clustvarsel)
library(mclust)
library(ggpubr)
library(svglite)
```

Import the dataframe created in the load_dataframe.rmd
```{r}
df_lm <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/df_lm.rds")
df_lmw <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/df_lmw.rds")
df_iso <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/df_iso.rds")

#df_NoLA <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/df_NoLA.rds")
```



## GMM

if only considering shape (because there are temperature as factor, so measurements not normalized by size / body length is not appropriate)
```{r}
df_NoLA <- df_iso %>% 
  dplyr::select(-c("ALO", "AW", "LTM", "LL", "sd_LL","sd_LTM","sd_ALO","sd_AW"))%>%
  drop_na() %>%
  filter(if_all(where(is.numeric),  ~ is.finite(.x) & .x != 0))
  #filter(if_all(where(is.numeric), is.finite))



measure_cols <- colnames(df_NoLA[,7:19])
#df_NoLA <-  df_NoLA  %>%  
#  mutate(across(all_of(measure_cols), 
#                ~ log(.x/exp(mean(log(.x)))), 
#                .names = "shape_ratio_{.col}"))

```

### remove weird individual
```{r}
length(unique(df_NoLA$sample_id))
df_NoLA <- df_NoLA  %>% filter(sample_id != "1-18-202510-dev-seq-A")
length(unique(df_NoLA$sample_id))
```

```{r}
# linear measurement
#df_ori <- df_NoLA[,7:19]
# log transformed linear measurement
#df_log <- log(df_NoLA[,7:19])
# measurement standardized by estimated volume
#df_ratio <- df_NoLA[,21:33]
# log ratio of measurement standardized by estimated volume
#df_log_shape_ratio <- log(df_NoLA[,21:33])
# log ratio by geometric means
#df_log_shape_ration_gm <- df_NoLA[,36:48]

# measurement standardized by estimated size (BL)
df_ratio <- df_NoLA[,20:32]
# log ratio of measurement standardized by estimated volume
df_log_shape_ratio <- log(df_NoLA[,20:32])
```

removing the measurements that occur in size estimation
```{r}
df_ratio <- df_ratio %>% 
  dplyr::select(-c("sd_BL"))

df_log_shape_ratio <- df_log_shape_ratio %>% 
  dplyr::select(-c("sd_BL"))
```

```{r}
#mod1 <- Mclust(df_ori,G=1:12)
#summary(mod1)
# mod1: 9 clusters

#mod2 <- Mclust(df_log,G=1:12)
#summary(mod2)
# mod2: 6 clusters

# measurement standardized by estimated size (BL)
mod3 <- Mclust(df_ratio,G=1:12)
summary(mod3)
# mod3: 4 cluster

mod4 <- Mclust(df_log_shape_ratio,G=1:12)
summary(mod4)
# mod4: 4 cluster

#mod5 <- Mclust(df_log_shape_ration_gm,G=1:12)
#summary(mod5)
# mod4: 5 cluster


```

```{r}
#p <- fviz_mclust(mod1,"classification",  geom = "point" ) +
#scale_shape_manual(values = c(16,0,17,8))
#scale_color_manual(values = palette3)+
#scale_fill_manual( values = palette3)
#p2 <- fviz_mclust(mod2,"classification",  geom = "point" ) +
#scale_shape_manual(values = c(16,0,17,8))
#scale_color_manual(values = palette3)+
#scale_fill_manual( values = palette3)
p3 <- fviz_mclust(mod3,"classification",  geom = "point" ) +
#scale_shape_manual(values = c(16,0,17,8))
scale_color_manual(values = palette5)+
scale_fill_manual( values = palette5)
p4 <- fviz_mclust(mod4,"classification",  geom = "point" ) +
#scale_shape_manual(values = c(16,0,17,8))
scale_color_manual(values = palette5)+
scale_fill_manual( values = palette5)
#p5 <- fviz_mclust(mod5,"classification",  geom = "point" ) +
#scale_shape_manual(values = c(16,0,17,8))
#scale_color_manual(values = palette3)+
#scale_fill_manual( values = palette3)
p3
p4
#pGMM <- ggarrange(p, p2, p3,p4, nrow = 2, ncol = 2,common.legend = TRUE)
ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/gmm_BL_standardized_cluster_result.svg", plot=p3)
#ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/gmm_BL_sd_log.svg", plot=p4)
#ggsave(file="D:/202603-dev-seq-analysis/fig/GMM_PCA.svg", plot=p3)
```



```{r}
#svglite("/Volumes/Yinghui/202603-dev-seq-analysis/fig/GMM_BIC_selected_model.svg", width = 7, height = 7)
#modelNames = c("VVE", "VEE","EEE", "VEV"),
plot(mod3, what = 'BIC')
#dev.off()
```

```{r}
min_idx <- which.min(pca$x[, 1])
df_gmm_export[min_idx,]

max_idx <- which.max(pca$x[, 2])
df_gmm_export[max_idx,]
```

```{r}

#df_NoLA$cluster <- as.factor(mod1$classification)
#df_NoLA$cluster_log <- as.factor(mod2$classification)
df_NoLA$cluster_shape <- as.factor(mod3$classification)
df_NoLA$cluster_shape_ratio <- as.factor(mod4$classification)
```


###Finding representing specimens

```{r}
n_clusters <- mod3$G
indices_of_exemplars <- c()

# 2. Loop through each cluster
for (i in 1:n_clusters) {
  
  # Get the mathematical centroid for this cluster
  center <- mod3$parameters$mean[, i]
  
  # Identify which rows belong to this cluster
  cluster_rows <- which(mod3$classification == i)
  
  # Calculate distance from each row in the cluster to the centroid
  # use the original data for the distance calculation
  distances <- apply(df_ratio[cluster_rows, , drop = FALSE], 1, function(x) {
    sum((x - center)^2)
  })
  
  # Find the index of the minimum distance
  exemplar_index <- cluster_rows[which.min(distances)]
  
  # Store it
  indices_of_exemplars[i] <- exemplar_index
}

# 3. Create a summary table of the representative specimens
representative_specimens <- df_NoLA[indices_of_exemplars, ]
print(representative_specimens)
```
### identify cluster composition
```{r}
library(RColorBrewer)
n <- 60
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
pie(rep(1,n), col=sample(col_vector, n))
```
individual composition for each cluster

```{r}
ggplot(df_NoLA, aes(x = factor(cluster_shape), fill = sample_id)) +
  geom_bar(position = "stack") +
  #scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = col_vector)+
  theme_classic()
```
```{r}
temp <- df_NoLA %>%
  group_by(cluster_shape) %>%
  summarise(n_samples = n_distinct(sample_id)) %>%
  ggplot(aes(x = factor(cluster_shape), y = n_samples)) +
  geom_col()  + 
  theme_classic() 

num <- df_NoLA %>%
  group_by(cluster_shape) %>%
  summarise(n_samples = n_distinct(sample_id)) 

num

temp
```

### identify contribution

```{r}
#ggsave(file="D:/202603-dev-seq-analysis/fig/GMM_PCA.svg", plot=p)
# red line: expected average contribution.
pca <- prcomp(scale(mod3$data))
fviz_contrib(pca, choice = "var", axes = 1, top = 10)
fviz_contrib(pca, choice = "var", axes = 2, top = 10)
fviz_contrib(pca, choice = "var", axes = 3, top = 10)
```

construct the plot after selecting the variables
```{r}
#out1 <- clustvarsel(df_ori)
#mod1_1 <-Mclust(df_ori[, out1$subset])
#summary(mod1_1)
#fviz_mclust(mod1_1,"classification",  geom = "point" ) 
```

```{r}
library(ggalluvial)


ggplot(df_NoLA, aes(x = BL, y = TABW, color = cluster_shape)) +
geom_point(size = 1) +
scale_color_manual(values = palette5)+
#geom_smooth()+
theme_classic()
```



```{r}
g_plot <- ggplot(df_NoLA,
aes(axis1 = stage, axis2 = cluster_shape, y = exp_day)) +
geom_alluvium(aes(fill = stage)) +
geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
theme_void() +
labs(x = "Classification Method",
y = "Number of images")
#scale_fill_manual(values = c("#343434","#3880ba","#a4cdee"))
g_plot
#ggsave(file="D:/202603-dev-seq-analysis/fig/ggalluvial.svg", plot=g_plot)
```

## Regression


```{r}
library("RRPP")

# Create a matrix of shape variables
shape_matrix<- df_iso %>% ungroup()%>%
  dplyr::select(-c("LL","LTM","ALO","AW","BL","sd_BL"))%>%
  dplyr::select(starts_with("sd_")) %>%
  drop_na() %>%
   filter(if_any(where(is.numeric), is.finite)) %>% 
  as.matrix()

shape_matrix <- df_iso %>% ungroup()%>%
  dplyr::select(-c("LL","LTM","ALO","AW"))%>%
  drop_na() %>%
  dplyr::select(7:19)%>%
   filter(if_any(where(is.numeric), is.finite)) %>% 
  as.matrix()

#shape_matrix

df_temp <- df_iso %>% ungroup()%>%
  dplyr::select(-c("LL","LTM","ALO","AW"))%>%
  drop_na()%>%
   filter(if_any(where(is.numeric), is.finite))

nrow(shape_matrix)
nrow(df_temp)

# Add the body length back to a clean modeling data frame
model_data <- rrpp.data.frame(
  BL = df_temp$BL,
  coords = shape_matrix,
  temp = as.factor(df_temp$temp),
  age = df_temp$exp_day,
  biorep = df_temp$biorep
)
#any(!is.finite(shape_matrix)) # Should be FALSE
#table(model_data$temp)
fit <- lm.rrpp(coords ~ log(BL)*temp*age, SS.type = "I", 
data = model_data, print.progress = FALSE, iter = 999)
summary(fit)
anova(fit)
```

```{r}
library(lme4)
library(lmerTest)
pca_shape <- prcomp(df_log_shape_ratio, scale. = TRUE)
summary(pca_shape)

pc1_scores <- pca_shape$x[, 1]

standard_df <- data.frame(
  pc1   = pc1_scores,
  vol = df_NoLA$vol,
  temp     = df_NoLA$temp,
  age      = df_NoLA$exp_day,
  biorep   = df_NoLA$biorep
)

# Run the model again
model <- lmer(pc1 ~ log(vol) * temp * age + (1 | biorep),
              data = standard_df)
summary(model)
anova(model)
```

```{r}
library(ggfortify)
 pca1 <- prcomp(df_NoLA[,21:33])
summary(pca1)
autoplot(pca1,data = df_NoLA, colour = 'stage',frame = TRUE, frame.type = 'norm')
```
```{r}

fviz_contrib(pca1, choice = "var", axes = 1, top = 10)
fviz_contrib(pca, choice = "var", axes = 2, top = 10)
fviz_contrib(pca, choice = "var", axes = 3, top = 10)
```

## Export the cluster data as dataframe
```{r}
#df_NoLA$cluster_raw_lm <- as.factor(mod1$classification)
#df_NoLA$cluster_log_lm <- as.factor(mod2$classification)
#df_NoLA$cluster_ratio_vol <- as.factor(mod3$classification)
#df_NoLA$cluster_shape_ratio_vol <- as.factor(mod4$classification)
#df_NoLA$cluster_shape_ratio_gm <- as.factor(mod5$classification)

df_gmm_export <- df_NoLA %>% ungroup() %>% mutate(cluster_shape = as.factor(mod3$classification))

#df_gmm_export <- df_iso %>%
#  left_join(df_gmm_export, by = c("sample_id" = "sample_id", "date_obj" = "date_obj")) 
```

```{r}
#saveRDS(df_NoLA, file = "/Volumes/Yinghui/202603-dev-seq-analysis/data/df_NoLA.rds")
saveRDS(df_gmm_export, file = "/Volumes/Yinghui/202603-dev-seq-analysis/data/df_gmm_export.rds")

```
