---
title: "202512_morph_lm"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "D:/202512-dev-seq-analyses/")

source("data_import.R")
source("PA_measurement.R")
source("TA_measurement.R")
source("OA_measurement.R")
```

```{r}
library("geomorph")
library("dplyr")
options(rgl.useNULL=TRUE)
library("rgl")
library("tidyverse")
library("gtools")
library(lubridate)
library(gridExtra)
library(factoextra)
library(ggplot2)
library(clustvarsel)
library(mclust)
library(ggpubr)
library(svglite)
```


# Import data file

Dir:"/Volumes/Yinghui/202512-dev-seq-analyses"

```{r}
# df <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds")
#df <- readRDS("D:/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds")

#files <- c("D:/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds", "D:/202603-dev-seq-analysis/data/202509-dev-seq-v6.rds")

files <- c("/Volumes/Yinghui/202603-dev-seq-analysis/data/202510-dev-seq-v6.rds", "/Volumes/Yinghui/202603-dev-seq-analysis/data/202509-dev-seq-v6.rds")

df <- lapply(files, readRDS) %>% bind_rows()
```



```{r}
# convert the lm to matrix for each file.

specs_list <- df %>%
  group_by(temp, biorep, ind, plane, file) %>%
  group_map(~ {
    id <- .y$file
    # Create a sub-list for this specific identifier
    spec <- list(
      matrix = as.matrix(.x[, c("X", "Y")]),
      metadata = as.list(.y)
    )
    
    setNames(list(spec), id)
  }) %>%
  flatten()

```

```{r}
# add lm 13 in the TA plane
specs_list<-add_lm13(specs_list)
```


# Plot setting
```{r}
palette1 <- c("#343434","#A3CEF1","#3880ba","#224f73","#000000","#CCCAC2","#837D72","#ede7da")
```

#linear measurements
```{r}
measurement_accumulator <- list()

for (id in names(specs_list)) {
  
  # 1. Extract components correctly
  spec   <- specs_list[[id]]
  mat    <- spec$matrix
  plane  <- spec$metadata$plane
  ind    <- spec$metadata$ind
  temp <- spec$metadata$temp
  biorep <- spec$metadata$biorep
  file <- spec$metadata$file
  measurer <- spec$metadata$measurer
  
 # print(file)
  # 2. Initialize a placeholder for this specific iteration's results
  res <- tibble()
  
 # print(file)
  # 3. Apply custom "Design" logic
  if (plane == "PA") {
    res <- tibble(
      measure_name = c("ALO", "AW", "WAE", "ATCL", "PABW"),
      value = c(ALO_fun(mat), AW_fun(mat), WAE_fun(mat), ATCL_fun(mat), PABW_fun(mat))
    )
    
  } else if (plane == "TA") {
    res <- tibble(
      measure_name = c("PL", "BL", "TCL", "LTM", "LL", "TABW", "AECL","TATB"),
      value = c(PL_fun(mat), BL_fun(mat), TCL_fun(mat), LTM_fun(mat), 
                LL_fun(mat), TABW_fun(mat), AECL_fun(mat), TATB_fun(mat))
    )
    
  } else if (plane == "OA") {
    res <- tibble(
      measure_name = c("TBL", "PW", "ATAC", "AEAC"),
      value = c(TBL_fun(mat), PW_fun(mat), ATAC_fun(mat), AEAC_fun(mat))
    )
  }
  
  # print(file)
  # print(row(mat))
  # print(file)
  # 4. Add metadata and store in accumulator
  # Only add if we actually generated results
  if (nrow(res) > 0) {
    measurement_accumulator[[id]] <- res %>%
      mutate(
        measurer = measurer,
        sample_id = str_extract(basename(file), "^\\d+"),
        ind = ind,
        plane = plane,
        temp = temp,
        biorep = biorep
      )
  }
}
# 5. Bind into one large dataframe
df_lm <- bind_rows(measurement_accumulator)

df_lm <- df_lm %>% mutate(
    # Convert from microns (or current unit) to mm
    value = value / 1000
  ) 
```

Removing the value of lobate and auricle when those structures are not present
standard:
ALO 0.07mm
AW 0.07mm
LTM 0.07mm
LL 0.07mm
```{r}
threshold <- 0.07

df_lm <- df_lm %>% mutate(across(
    c(value), 
    ~ if_else(measure_name %in% c("ALO", "AW", "LTM", "LL") & .x < threshold, NA_real_, .x)
  ))
  
```

convert the sample_id to date
```{r}
df_lm <- df_lm %>%
  mutate(date_obj = parse_date_time(paste0(sample_id, "2026"), orders = "mdy")) %>%
  group_by(biorep) %>%
  mutate(exp_day = as.numeric(difftime(date_obj, min(date_obj), units = "days"))) %>%
  ungroup() %>%
  dplyr::select(-c(date_obj))
```
change the sample id to individual
```{r}
df_lm <- df_lm  %>%
  mutate(sample_id = paste(ind, temp, biorep, sep = "-"))
```

## allometric dev visualisation

### obtain the natural log of the value
Common practice when studying allometric development because it is easier to deal with proportional variation.

```{r}
#create a new column with log value
df_lm <- df_lm %>% mutate(log_value = log(value))
```

###Plotting for the shape versus size LMM study

centroid size is recalculated as a Solid semi-ellipsoid around z-axis (without lobe)

volume: (2*pi*a*b*c)/3

a, b = TABW/2, PABW/2, the principal semi-axes of the base ellipse
c = BL,the principal z-semi-axe from the center of base ellipse

```{r}
df_lmw <- df_lm %>%
  dplyr::select(-c(plane, log_value)) %>%
  group_by(sample_id, ind) %>%
  pivot_wider(
    names_from = measure_name,
    values_from = value
  ) 

df_lmw <- df_lmw %>% 
  mutate(centroid = (2 *(TABW/2) * (PABW/2)*BL )/3)


measure_cols <- colnames(df_lmw[,6:22])
df_iso <-  df_lmw  %>%  
  mutate(across(all_of(measure_cols), 
                ~ log(.x/centroid), 
                .names = "shape_{.col}"))

```


```{r}
p1 <-ggplot(df_lmw, aes(x = log(centroid), y = log(TBL), colour = temp)) +
    geom_point(size = 2, alpha = 0.8) +
    # Add the dashed regression line
    geom_smooth(method = "lm", formula = y ~ x, color = "black") +
    theme_classic() +
    labs(x = "Log Centroid size", y = "Log of the tentacle bulb length (mm)") +
    scale_color_manual(values = c( "#00B2CA", "#F79256")) +
    theme(legend.position = "right")
#+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))
p4 <-ggplot(df_lmw, aes(x = log(centroid), y = log(TABW), colour = temp)) +
    geom_point(size = 2, alpha = 0.8) +
    # Add the dashed regression line
    geom_smooth(method = "lm", formula = y ~ x, color = "black") +
    theme_classic() +
    labs(x = "Log Centroid size", y = "Log of the body width (mm)") +
    scale_color_manual(values = c( "#00B2CA", "#F79256")) +
    theme(legend.position = "right")
#+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))
p2 <- ggplot(df_lmw, aes(x = log(centroid), y = log(LL), colour = temp)) +
    geom_point(size = 2, alpha = 0.8) +
    # Add the dashed regression line
    geom_smooth(method = "lm", formula = y ~ x, color = "black") +
    theme_classic() +
    labs(x = "Log Centroid size", y = "Log of the lobe side length (mm)") +
    scale_color_manual(values = c("#00B2CA", "#F79256")) +
    theme(legend.position = "right")
#+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))
p3 <- ggplot(df_lmw, aes(x = log(centroid), y = log(ALO), colour = temp)) +
    geom_point(size = 2, alpha = 0.8) +
    # Add the dashed regression line
    geom_smooth(method = "lm", formula = y ~ x, color = "black") +
    theme_classic() +
    labs(x = "Log Centroid size", y = "Log of the auricle length (mm)") +
    scale_color_manual(values = c("#00B2CA", "#F79256")) +
    theme(legend.position = "right")
#+ stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")))

p <- ggarrange(p1,p4, p2,p3, nrow = 2, ncol = 2,common.legend = TRUE)
#ggsave(file="/Volumes/Yinghui/202512-dev-seq-analyses/fig/p.svg", plot=p)
#ggsave(file="D:/202603-dev-seq-analysis/fig/p1.svg", plot=p)

```

#### With all variables:

```{r}
measure_cols <- colnames(df_lmw[,6:22])
df_iso <-  df_lmw  %>%  
  mutate(across(all_of(measure_cols), 
                ~ log(.x/centroid), 
                .names = "shape_{.col}"))
```


```{r}
tbl_values <- df_lm$value[df_lm$measure_name == "TBL"]
boxplot(as.numeric(tbl_values))

```