---
title: "lm_WGCNA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

```{r}
library("dplyr")
options(rgl.useNULL=TRUE)
library("rgl")
library("tidyverse")
library("gtools")
library(lubridate)
library(gridExtra)
library(factoextra)
library(ggplot2)
library(clustvarsel)
library(ggpubr)
library(svglite)
library(MASS)

# WGCNA
#install.packages('BiocManager')
#library(BiocManager)
#BiocManager::install('WGCNA')
#BiocManager::install('flashClust')
library(WGCNA)
library(flashClust)
library(curl)
library(ggcorrplot)
```

Import the dataframe created in the load_dataframe.rmd
```{r}
df_lm <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/df_lm.rds")
df_lmw <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/df_lmw.rds")
df_iso <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/df_iso.rds")
#df_NoLA_6.4 <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/df_NoLA.rds")
#df_NoLA <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/df_NoLA.rds")
df_gmm_export <- readRDS("/Volumes/Yinghui/202603-dev-seq-analysis/data/df_gmm_export.rds")

```

# Tutorial
```{r}
d <- curl('https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/bioanth-stats/module-F21-Group1/FemaleLiver-Data/LiverFemale3600.csv')
liver.data <- read.csv(file = d, stringsAsFactors = FALSE, header = TRUE)
head(liver.data)
```
## Data Cleaning and Removing Outliers
```{r}
# remove extra trait
expression.data <- liver.data[,-c(1:8)] #removing variables not holding expression data

expression.data <- as.data.frame(t(expression.data)) #transforming the data.frame so columns now represent genes and rows represent samples
names(expression.data) <- liver.data$substanceBXH
#renaming the columns so we don't lose the gene IDs
```
### identify outlier
The WGCNA package has a built in function to identify outlier genes called goodSampleGenes(). The function checks the data and returns a list object of samples and genes that pass its filtering criteria. You can adjust how strict the filtering process is by changing the default of the arguments (which can be found under the goodSampleGene documentation, ?goodSampleGenes).
```{r}
gsg <-goodSamplesGenes(expression.data)
summary(gsg)
```
Identifying outlier samples
You can identify outlier samples by using hierarchical clustering. It is probably beneficial to review the clustering module as clustering is used several times within this pathway.

```{r}
sampleTree <- hclust(dist(expression.data), method = "average") #Clustering samples based on distance 

#Setting the graphical parameters
par(cex = 0.6);
par(mar = c(0,4,2,0))

#Plotting the cluster dendrogram
plot(sampleTree, main = "Sample clustering to detect outliers", sub="", xlab="", cex.lab = 1.5,
cex.axis = 1.5, cex.main = 2)

#Setting the graphical parameters
par(cex = 0.6);
par(mar = c(0,4,2,0))
plot(sampleTree, main = "Sample clustering to detect outliers", sub="", xlab="", cex.lab = 1.5,
cex.axis = 1.5, cex.main = 2)
#draw on line to show cutoff height
abline(h = 15, col = "red");
```
```{r}
cut.sampleTree <- cutreeStatic(sampleTree, cutHeight = 15, minSize = 10) #returns numeric vector
#Remove outlier
expression.data <- expression.data[cut.sampleTree==1, ]
```

## Network Construction
```{r}

```

# Correlation analysis
```{r}
# clean data
# from df_iso, getting rid of the data that has NA in measurements that is not lobe or auricle
df_clean <- df_iso %>% drop_na(-c(ALO, AW, LL, LTM, sd_ALO, sd_AW, sd_LL, sd_LTM))

# turn all lobe and auricle measurements to 0
df_clean[is.na(df_clean)] <- 0

# extract measurements
#df_corr <- df_clean[7:23]
df_corr <- df_clean[24:40]

```

```{r}
df_corr <- df_corr[-11]
# plot correlation
corr <- cor(df_corr, method = "spearman")
corr_heatmap <- ggcorrplot(corr,hc.order = TRUE, lab = TRUE)
ggsave(file="/Volumes/Yinghui/202603-dev-seq-analysis/fig/corr_heatmap.svg", plot=corr_heatmap)
```